---
title: "Portable Polarization: Partisan Identity, Expressive Attitudes, and COVID-19 Health Behavior"
format: docx
bibliography: references.bib
filters:
  - wordcount.lua
editor:
  markdown:
    wrap: sentence
---

Christopher Soria^1^, Audrey M. Dorélien^2^, Ayesha Mahmud^1^, Dennis Feehan^1^

^1^Department of Demography, University of California, Berkeley
^2^Department of Sociology, University of Washington

```{r}
#| label: setup
#| include: false

library(tidyverse)

# Load pre-calculated paper statistics
stats_path <- "../../out/tables/paper_statistics.rds"
if (!file.exists(stats_path)) stop("Run code/export_paper_statistics.R first.")
paper_stats <- readRDS(stats_path)
list2env(paper_stats, envir = environment())

# Load wave-level adjusted gaps (contacts and masks)
adj_gaps <- read_csv("../../out/tables/adjusted_gaps_by_wave.csv",
                     show_col_types = FALSE)

contacts_adj <- adj_gaps |> filter(outcome == "contacts")
masks_adj    <- adj_gaps |> filter(outcome == "masks")

# Peak contact gap: wave with largest Rep-Dem adjusted gap
peak_contact_wave <- contacts_adj |> slice_max(gap_rep_dem, n = 1)
peak_contact_gap  <- peak_contact_wave$gap_rep_dem
peak_contact_wavenum <- peak_contact_wave$wave   # wave 3 = Sep '20

# Smallest contact gap: wave 5 (winter 2020-21 surge)
w5_contact_gap <- contacts_adj |> filter(wave == 5) |> pull(gap_rep_dem)

# Mask gaps at wave 1 and wave 6 (Dem - Rep, so positive = Dem higher)
w1_mask_gap <- masks_adj |> filter(wave == 1) |> pull(gap_ind_dem) |> abs()
w1_mask_gap <- masks_adj |> filter(wave == 1) |>
  mutate(gap = Democrat - Republican) |> pull(gap)
w6_mask_gap <- masks_adj |> filter(wave == 6) |>
  mutate(gap = Democrat - Republican) |> pull(gap)

# Load regional adjusted gaps
reg_gaps <- read_csv("../../out/tables/adj_gaps_by_region.csv",
                     show_col_types = FALSE)

contact_gap_min <- min(reg_gaps$contacts_gap_rd)
contact_gap_max <- max(reg_gaps$contacts_gap_rd)
mask_gap_min    <- min(reg_gaps$mask_gap_dr)
mask_gap_max    <- max(reg_gaps$mask_gap_dr)

# Load fully-adjusted partisan gaps (Republican vs Democrat, Model 3)
adj_party_gaps <- read_csv("../../out/tables/adjusted_partisan_gaps.csv",
                            show_col_types = FALSE)
contact_adj_coef <- adj_party_gaps |> filter(outcome == "contacts_total") |> pull(estimate)
contact_adj_se   <- adj_party_gaps |> filter(outcome == "contacts_total") |> pull(std_error)
mask_adj_coef    <- adj_party_gaps |> filter(outcome == "mask_usage") |> pull(estimate)
mask_adj_se      <- adj_party_gaps |> filter(outcome == "mask_usage") |> pull(std_error)

# Load wave-level gaps with mortality context (for Figure 4)
gaps_by_wave <- read_csv("../../out/tables/gaps_by_wave.csv",
                         show_col_types = FALSE) |>
  mutate(
    wave_label = case_when(
      wave == 1 ~ "Apr '20",
      wave == 2 ~ "Jun '20",
      wave == 3 ~ "Sep '20",
      wave == 4 ~ "Dec '20",
      wave == 5 ~ "Feb '21",
      wave == 6 ~ "May '21"
    ),
    wave_label = factor(wave_label,
                        levels = c("Apr '20", "Jun '20", "Sep '20",
                                   "Dec '20", "Feb '21", "May '21"))
  )
```

**Keywords:** Political polarization; Partisan identity; COVID-19; Health behavior; Social contact patterns

**Status and ethics:** This manuscript reports original research and is being prepared for submission; it has not been published or submitted elsewhere. The study uses survey data originally collected by the authors during the COVID-19 pandemic; as a secondary analysis of de-identified data, this work does not require additional Institutional Review Board review.

## Abstract

Health behaviors during the COVID-19 pandemic differed by political party affiliation [@gadarian_partisanship_2021; @allcott_polarization_2020; @clinton_partisan_2021].
Building on prior work showing these partisan gaps can alter modeled epidemic dynamics [@soria_2026], we examine how the polarization persists across contexts and over time, and whether mask and contact gaps follow different patterns.
Using six cross-sectional waves of the Berkeley Interpersonal Contact Study (BICS), we test whether they reduce to mask-mandate policy, demographic composition, geographic context, or perceived severity — distinctions that matter for transmission modeling and for understanding when health issues lock into partisan polarization more broadly.

BICS — one of the few US nationally representative datasets combining detailed face-to-face contact diaries with individual partisan identity — spans April 2020 to May 2021 (N = `r format(n_total, big.mark = ",")`), recording daily face-to-face contacts and mask usage.
We linked individual survey responses to county-level COVID-19 incidence and mortality rates, congressional district partisan composition, and county mask mandate status.
We used progressively nested regression models with subgroup analyses.

Across all six waves, Republicans reported substantially more daily contacts and lower rates of mask usage and vaccine acceptance than Democrats.

These gaps did not move in tandem with pandemic severity. The mask gap widened across the study period (from `r sprintf("%.1f", w1_mask_gap)` to `r sprintf("%.1f", w6_mask_gap)` percentage points, with a brief narrowing in September 2020), and Republicans reported their lowest concern about COVID at the winter 2020–21 mortality peak. The contact gap, in contrast, narrowed substantially during that same surge, shrinking to `r sprintf("%.1f", w5_contact_gap)` contacts/day in February 2021.

The mask gap was largely insensitive to political and policy context. Republicans in Democrat-majority districts wore masks at higher rates than co-partisans in Republican-majority areas, but Democrats shifted upward in those contexts as well, and the partisan gap persisted within every mask-mandate regime, leaving polarization intact across both political and policy environments.

The partisan contact gap diminished sharply in some subgroups (approaching zero among non-metropolitan and non-White respondents), while a partisan mask gap was present in every subgroup we examined. Despite this heterogeneity, partisan identity remained a statistically significant predictor in fully adjusted models: relative to Democrats, Republicans reported `r sprintf("%.2f", contact_adj_coef)` more contacts per day (SE = `r sprintf("%.2f", contact_adj_se)`, p < .001) and `r sprintf("%.2f", abs(mask_adj_coef))` percentage points lower mask usage (SE = `r sprintf("%.2f", mask_adj_se)`, p < .001).

```{r}
#| label: fig-partisan-gaps
#| fig-cap: "Partisan Gaps in Health Behaviors Across Survey Waves. Lines show the Republican–Democrat difference in daily contacts (Panel A) and the Democrat–Republican difference in mask usage (Panel B). Gray bars show local mortality rates (deaths per 100,000) during each wave."
#| fig-height: 5
#| fig-width: 10
#| echo: false
#| warning: false
#| message: false

library(patchwork)
library(cowplot)

max_mort <- max(gaps_by_wave$mean_mort_rate)
max_contact_gap <- max(gaps_by_wave$contacts_gap)
max_mask_gap <- max(gaps_by_wave$mask_gap)

contact_scale <- max_contact_gap / max_mort * 1.3
mask_scale    <- max_mask_gap / max_mort * 1.3

p_contact_gap <- ggplot(gaps_by_wave, aes(x = wave_label)) +
  geom_col(aes(y = mean_mort_rate * contact_scale),
           fill = "gray70", alpha = 0.5, width = 0.7) +
  geom_line(aes(y = contacts_gap, group = 1), color = "#7570b3", linewidth = 1.2) +
  geom_point(aes(y = contacts_gap), color = "#7570b3", size = 4) +
  geom_text(aes(y = contacts_gap, label = sprintf("%.1f", contacts_gap)),
            vjust = -1.2, size = 4, fontface = "bold") +
  geom_hline(yintercept = 0, linetype = "dashed", color = "gray50") +
  scale_y_continuous(
    name = "R-D Gap (contacts/day)",
    sec.axis = sec_axis(~ . / contact_scale, name = "Mortality (per 100k)"),
    expand = expansion(mult = c(0.05, 0.2))
  ) +
  labs(x = NULL, title = "A. Contact Gap (Republican - Democrat)") +
  theme_cowplot() +
  theme(
    plot.title = element_text(face = "bold", size = 15),
    axis.text.x = element_text(angle = 45, hjust = 1),
    axis.title.y.right = element_text(color = "gray50")
  )

p_mask_gap <- ggplot(gaps_by_wave, aes(x = wave_label)) +
  geom_col(aes(y = mean_mort_rate * mask_scale),
           fill = "gray70", alpha = 0.5, width = 0.7) +
  geom_line(aes(y = mask_gap, group = 1), color = "#1b9e77", linewidth = 1.2) +
  geom_point(aes(y = mask_gap), color = "#1b9e77", size = 4) +
  geom_text(aes(y = mask_gap, label = sprintf("%.1f", mask_gap)),
            vjust = -1.2, size = 4, fontface = "bold") +
  geom_hline(yintercept = 0, linetype = "dashed", color = "gray50") +
  scale_y_continuous(
    name = "D-R Gap (percentage points)",
    sec.axis = sec_axis(~ . / mask_scale, name = "Mortality (per 100k)"),
    expand = expansion(mult = c(0.05, 0.2))
  ) +
  labs(x = NULL, title = "B. Mask Gap (Democrat - Republican)") +
  theme_cowplot() +
  theme(
    plot.title = element_text(face = "bold", size = 15),
    axis.text.x = element_text(angle = 45, hjust = 1),
    axis.title.y.right = element_text(color = "gray50")
  )

p_contact_gap + p_mask_gap +
  plot_annotation(caption = "Note: Gray background bars show local mortality rates during each survey wave (right axis).")
```

These patterns extend prior work showing that partisan identity structures publicly observable health behaviors more strongly than less-visible ones [@baxter-king_how_2022].
Contact gaps were smallest at peak mortality, while the mask gap widened despite fluctuations in conditions, consistent with masking as a publicly legible group marker subject to identity costs [@goffman_stigma_1963].
Republicans' lowest concern coincided with peak mortality, consistent with expressive responding shaping stated concern [@clinton_partisan_2021].
These gaps rival those by race, gender, and education and persist — particularly for masks — across policy and contextual controls, pointing to partisan identity as a robust correlate of expressive, more than instrumental, COVID-19 health behaviors, with implications for health inequality, partisan polarization, and the broader question of which health behaviors become locked into political identity in polarized contexts.
