---
title: "differential_descriptives"
author: "Christopher Soria"
date: "2024-08-08"
output: html_document
---

# =============================================================================
# TABLE OF CONTENTS
# -----------------------------------------------------------------------------
# 1. SETUP AND CONFIGURATION
# 2. CONTACTS DIFFERENCES
# 3. MASK USAGE DIFFERENCES
# 4. VACCINATION DIFFERENCES
# 5. COMBINED PLOTS
# 6. WAVE-LEVEL DIFFERENCES
# =============================================================================

# =============================================================================
# 1. SETUP AND CONFIGURATION
# =============================================================================

```{r setup, message=FALSE, warning=FALSE}
library(emmeans)
library(cowplot)
library(tidyverse)

# Source helper functions
source("plot_helper_functions/difference_helpers.R")

# Load Data
if (!exists("bics_zip_features")) {
  if (file.exists("../data/BICS_ZIP_Features.csv")) {
    bics_zip_features <- read_csv("../data/BICS_ZIP_Features.csv")
  } else {
    stop("Data file ../data/BICS_ZIP_Features.csv not found. Please run 00-Prep_BICS_Data.Rmd first.")
  }
}

# Create party-specific dataframes
RD_df <- bics_zip_features %>% filter(political_party %in% c("Republican", "Democrat"))
ID_df <- bics_zip_features %>% filter(political_party %in% c("Independent", "Democrat"))
D_df <- bics_zip_features %>% filter(political_party == "Democrat")
R_df <- bics_zip_features %>% filter(political_party == "Republican")

# Age-specific dataframes
bics_65_plus <- bics_zip_features %>% filter(age >= 65)
RD_65_plus <- RD_df %>% filter(age >= 65)
ID_65_plus <- ID_df %>% filter(age >= 65)
```

# =============================================================================
# 2. CONTACTS DIFFERENCES
# =============================================================================


## contacts
```{r contacts}
# Run demographic models and generate contrasts
combined_contrasts <- run_demographic_models(
  data = bics_zip_features,
  outcome_var = "num_cc",
  data_RD = RD_df,
  data_ID = ID_df
)

print(combined_contrasts)

# Plot contrasts
pred_contacts <- plot_demographic_contrasts(
  combined_contrasts,
  title = "Difference in total number of contacts"
)

ggsave("contrast_contacts.png", plot = pred_contacts, path = "../out/plots", width = 12, height = 8, units = "in")

pred_contacts
```
### contacts (old)
```{r}
bi_republican <- lm(num_cc ~ republican, data = RD_65_plus, weights = weight_party_raked)
bi_independent <- lm(num_cc ~ Independent, data = ID_65_plus, weights = weight_party_raked)
bi_race <- lm(num_cc ~ white, data = bics_65_plus, weights = weight_party_raked)
bi_gender <- lm(num_cc ~ male, data = bics_65_plus, weights = weight_party_raked)
bi_concern <- lm(num_cc ~ strong_concern_inverse, data = bics_65_plus, weights = weight_party_raked)
bi_college <- lm(num_cc ~ r_college_grad, data = bics_65_plus, weights = weight_party_raked)

emm_race <- emmeans(bi_race, specs = pairwise ~ white)
emm_gender <- emmeans(bi_gender, specs = pairwise ~ male)
emm_republican <- emmeans(bi_republican, specs = pairwise ~ republican)
emm_independent <- emmeans(bi_independent, specs = pairwise ~ Independent)
emm_concern <- emmeans(bi_concern, specs = pairwise ~ strong_concern_inverse)
emm_college <- emmeans(bi_college, specs = pairwise ~ r_college_grad)

contr_race <- contrast(emm_race, "pairwise")
contr_gender <- contrast(emm_gender, "pairwise")
contr_republican <- contrast(emm_republican, "pairwise")
contr_independent <- contrast(emm_independent, "pairwise")
contr_college <- contrast(emm_college, "pairwise")

df_race_contrasts <- summary(contr_race) %>% as.data.frame()
df_gender_contrasts <- summary(contr_gender) %>% as.data.frame()
df_republican_contrasts <- summary(contr_republican) %>% as.data.frame()
df_independent_contrasts <- summary(contr_independent) %>% as.data.frame()
df_college_contrasts <- summary(contr_college) %>% as.data.frame()

# Combine all emm_* data frames into one
combined_contrasts <- bind_rows(df_independent_contrasts, df_republican_contrasts, df_college_contrasts, df_race_contrasts, df_gender_contrasts)
combined_contrasts <- as.data.frame(combined_contrasts)

combined_contrasts$contrast <- factor(combined_contrasts$contrast, levels = unique(combined_contrasts$contrast))

print(combined_contrasts)


pred_contacts_65 <- ggplot(combined_contrasts, aes(x = contrast, y = -1 * estimate, fill = contrast)) +
  geom_bar(stat = "identity", position = position_dodge(), width = 0.7, color = "black",  alpha = 0.7) +  # Draw bars
  geom_errorbar(aes(ymin = -1 * (estimate + SE), ymax = -1 * (estimate - SE)), 
                position = position_dodge(0.7), width = 0.25) +  # Add error bars
  labs(x = "", y = "", title = "Difference in total number of contacts") +
  scale_x_discrete(labels = c("Independent-\nDemocrat", "Republican-\nDemocrat", "College-\nNo-College", "White-\nNon-White", "Male-\nNon-Male")) +
  scale_fill_manual(values = c("Independent0 - Independent1" = "#6db753", "republican0 - republican1" = "#c50000",
                               "white0 - white1" = "gray", "male0 - male1" = "gray", "r_college_grad0 - r_college_grad1" = "gray")) +
  theme_cowplot() + 
  theme(legend.position = "none",
        plot.title = element_text(size = 20),
        panel.grid.minor.y = element_blank(),
        panel.grid.major.x = element_blank(),
        panel.grid.major.y = element_line(color = "grey90", size = 0.5),
        axis.title.x = element_text(size = 18),
        axis.text.x = element_text(size = 18),
        axis.text.y = element_text(size = 22),
        axis.title.y = element_text(size = 16),
        legend.title = element_text(size = 12),
        legend.text = element_text(size = 10))

ggsave("contrast_contact_65_plus.png", plot = pred_contacts_65, path = "../out/plots", width = 12, height = 8, units = "in")

pred_contacts_65

rm(emm_race, emm_gender, emm_age, emm_party, combined_contrasts, df_race_contrasts, df_gender_contrasts, df_age_contrasts, df_party_contrasts, contr_age, contr_gender, contr_party, contr_race, bi_age, bi_gender, bi_party, bi_race)
```


# =============================================================================
# 3. MASK USAGE DIFFERENCES
# =============================================================================

## mask usage
```{r masks}
# Run demographic models and generate contrasts
combined_contrasts_masks <- run_demographic_models(
  data = bics_zip_features,
  outcome_var = "Norm_Masks_Used",
  data_RD = RD_df,
  data_ID = ID_df
)

print(combined_contrasts_masks)

# Plot contrasts
pred_mask <- plot_demographic_contrasts(
  combined_contrasts_masks,
  title = "Difference in Percent of Contacts with a Mask"
) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "black")

ggsave("contrast_masks.png", plot = pred_mask, path = "../out/plots", width = 12, height = 8, units = "in")

pred_mask
```
### mask usage (old)
```{r}
bi_republican <- lm(Norm_Masks_Used ~ republican, data = RD_65_plus, weights = weight_party_raked)
bi_independent <- lm(Norm_Masks_Used ~ Independent, data = ID_65_plus, weights = weight_party_raked)
bi_race <- lm(Norm_Masks_Used ~ white, data = bics_65_plus, weights = weight_party_raked)
bi_gender <- lm(Norm_Masks_Used ~ male, data = bics_65_plus, weights = weight_party_raked)
bi_concern <- lm(Norm_Masks_Used ~ strong_concern_inverse, data = bics_65_plus, weights = weight_party_raked)
bi_college <- lm(Norm_Masks_Used ~ r_college_grad, data = bics_65_plus, weights = weight_party_raked)

emm_race <- emmeans(bi_race, specs = pairwise ~ white)
emm_gender <- emmeans(bi_gender, specs = pairwise ~ male)
emm_republican <- emmeans(bi_republican, specs = pairwise ~ republican)
emm_independent <- emmeans(bi_independent, specs = pairwise ~ Independent)
emm_concern <- emmeans(bi_concern, specs = pairwise ~ strong_concern_inverse)
emm_college <- emmeans(bi_college, specs = pairwise ~ r_college_grad)

contr_race <- contrast(emm_race, "pairwise")
contr_gender <- contrast(emm_gender, "pairwise")
contr_republican <- contrast(emm_republican, "pairwise")
contr_independent <- contrast(emm_independent, "pairwise")
contr_college <- contrast(emm_college, "pairwise")

df_race_contrasts <- summary(contr_race) %>% as.data.frame()
df_gender_contrasts <- summary(contr_gender) %>% as.data.frame()
df_republican_contrasts <- summary(contr_republican) %>% as.data.frame()
df_independent_contrasts <- summary(contr_independent) %>% as.data.frame()
df_college_contrasts <- summary(contr_college) %>% as.data.frame()

combined_contrasts <- bind_rows(df_independent_contrasts, df_republican_contrasts, df_college_contrasts, df_race_contrasts, df_gender_contrasts)
combined_contrasts <- as.data.frame(combined_contrasts)

combined_contrasts$contrast <- factor(combined_contrasts$contrast, levels = unique(combined_contrasts$contrast))

print(combined_contrasts)

pred_mask_65 <- ggplot(combined_contrasts, aes(x = contrast, y = -1 * estimate, fill = contrast)) +
  geom_bar(stat = "identity", position = position_dodge(), width = 0.7, color = "black",  alpha = 0.7) +  # Draw bars
  geom_errorbar(aes(ymin = -1 * (estimate + SE), ymax = -1 * (estimate - SE)), 
                position = position_dodge(0.7), width = 0.25) +  # Add error bars
  labs(x = "", y = "", title = "Difference in Percent of Contacts with a Mask") +
  scale_x_discrete(labels = c("Independent-\nDemocrat", "Republican-\nDemocrat", "College-\nNo-College", "White-\nNon-White", "Male-\nNon-Male")) +
  scale_fill_manual(values = c("Independent0 - Independent1" = "#6db753", "republican0 - republican1" = "#c50000",
                               "white0 - white1" = "gray", "male0 - male1" = "gray", "r_college_grad0 - r_college_grad1" = "gray")) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "black") + 
  theme_cowplot() +
  theme(legend.position = "none",
        plot.title = element_text(size = 20),
        panel.grid.minor.y = element_blank(),
        panel.grid.major.x = element_blank(),
        panel.grid.major.y = element_line(color = "grey90", size = 0.5),
        axis.title.x = element_text(size = 18),
        axis.text.x = element_text(size = 18),
        axis.text.y = element_text(size = 22),
        axis.title.y = element_text(size = 16),
        legend.title = element_text(size = 12),
        legend.text = element_text(size = 10))

ggsave("contrast_masks_65_plus.png", plot = pred_mask_65, path = "../out/plots", width = 12, height = 8, units = "in")

pred_mask_65

rm(emm_race, emm_gender, emm_age, emm_party, combined_contrasts, df_race_contrasts, df_gender_contrasts, df_age_contrasts, df_party_contrasts, contr_age, contr_gender, contr_party, contr_race, bi_age, bi_gender, bi_party, bi_race)
```


# =============================================================================
# 4. VACCINATION DIFFERENCES
# =============================================================================

# vaccination
```{r vaccines}
# Run demographic models and generate contrasts
combined_contrasts_vaccines <- run_demographic_models(
  data = bics_zip_features,
  outcome_var = "Vaccinated",
  data_RD = RD_df,
  data_ID = ID_df
)

print(combined_contrasts_vaccines)

# Plot contrasts
pred_vaccines <- plot_demographic_contrasts(
  combined_contrasts_vaccines,
  title = "Difference in Proportion Vaccinated"
) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "black")

ggsave("contrast_vaccines.png", plot = pred_vaccines, path = "../out/plots", width = 12, height = 8, units = "in")

pred_vaccines
```
### vaccination (old)
```{r}
bi_republican <- lm(Vaccinated ~ republican, data = RD_65_plus, weights = weight_party_raked)
bi_independent <- lm(Vaccinated ~ Independent, data = ID_65_plus, weights = weight_party_raked)
bi_race <- lm(Vaccinated ~ white, data = bics_65_plus, weights = weight_party_raked)
bi_gender <- lm(Vaccinated ~ male, data = bics_65_plus, weights = weight_party_raked)
bi_concern <- lm(Vaccinated ~ strong_concern_inverse, data = bics_65_plus, weights = weight_party_raked)
bi_college <- lm(Vaccinated ~ r_college_grad, data = bics_65_plus, weights = weight_party_raked)

emm_race <- emmeans(bi_race, specs = pairwise ~ white)
emm_gender <- emmeans(bi_gender, specs = pairwise ~ male)
emm_republican <- emmeans(bi_republican, specs = pairwise ~ republican)
emm_independent <- emmeans(bi_independent, specs = pairwise ~ Independent)
emm_concern <- emmeans(bi_concern, specs = pairwise ~ strong_concern_inverse)
emm_college <- emmeans(bi_college, specs = pairwise ~ r_college_grad)

contr_race <- contrast(emm_race, "pairwise")
contr_gender <- contrast(emm_gender, "pairwise")
contr_republican <- contrast(emm_republican, "pairwise")
contr_independent <- contrast(emm_independent, "pairwise")
contr_college <- contrast(emm_college, "pairwise")

df_race_contrasts <- summary(contr_race) %>% as.data.frame()
df_gender_contrasts <- summary(contr_gender) %>% as.data.frame()
df_republican_contrasts <- summary(contr_republican) %>% as.data.frame()
df_independent_contrasts <- summary(contr_independent) %>% as.data.frame()
df_college_contrasts <- summary(contr_college) %>% as.data.frame()

combined_contrasts <- bind_rows(df_independent_contrasts, df_republican_contrasts, df_college_contrasts, df_race_contrasts, df_gender_contrasts)
combined_contrasts <- as.data.frame(combined_contrasts)

combined_contrasts$contrast <- factor(combined_contrasts$contrast, levels = unique(combined_contrasts$contrast))

print(combined_contrasts)

pred_vaccines_65 <- ggplot(combined_contrasts, aes(x = contrast, y = -1 * estimate, fill = contrast)) +
  geom_bar(stat = "identity", position = position_dodge(), width = 0.7, color = "black",  alpha = 0.7) +  # Draw bars
  geom_errorbar(aes(ymin = -1 * (estimate + SE), ymax = -1 * (estimate - SE)), 
                position = position_dodge(0.7), width = 0.25) +  # Add error bars
  labs(x = "", y = "", title = "Difference in Proportion Vaccinated") +
  scale_x_discrete(labels = c("Independent-\nDemocrat", "Republican-\nDemocrat", "College-\nNo-College", "White-\nNon-White", "Male-\nNon-Male")) +
  scale_fill_manual(values = c("Independent0 - Independent1" = "#6db753", "republican0 - republican1" = "#c50000",
                               "white0 - white1" = "gray", "male0 - male1" = "gray", "r_college_grad0 - r_college_grad1" = "gray")) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "black") + 
  theme_cowplot() +
  theme(legend.position = "none",
        plot.title = element_text(size = 20),
        panel.grid.minor.y = element_blank(),
        panel.grid.major.x = element_blank(),
        panel.grid.major.y = element_line(color = "grey90", size = 0.5),
        axis.title.x = element_text(size = 18),
        axis.text.x = element_text(size = 18),
        axis.text.y = element_text(size = 22),
        axis.title.y = element_text(size = 16),
        legend.title = element_text(size = 12),
        legend.text = element_text(size = 10))


ggsave("contrast_vaccines_65_plus.png", plot = pred_vaccines_65, path = "../out/plots", width = 12, height = 8, units = "in")

pred_vaccines_65

rm(emm_race, emm_gender, emm_age, emm_party, combined_contrasts, df_race_contrasts, df_gender_contrasts, df_age_contrasts, df_party_contrasts, contr_age, contr_gender, contr_party, contr_race, bi_age, bi_gender, bi_party, bi_race)
```


# =============================================================================
# 5. COMBINED PLOTS
# =============================================================================

## merging differences plots
```{r}
combined_plot <- plot_grid(pred_contacts, pred_mask, pred_vaccines, ncol = 1,labels = c("A", "B", "C"))

combined_plot <- plot_grid(
  ggdraw() + draw_label("", size = 16, hjust = 0, x = 0.01),
  combined_plot,
  ncol = 1, 
  rel_heights = c(0.4, 10) 
)

combined_plot

ggsave("combined_contrasts.png", plot = combined_plot, path = "../out/plots", width = 12, height = 12, units = "in")

rm(combined_plot, pred_contacts, pred_mask, pred_vaccines)
```
## mergind differences plots (old)
```{r}
combined_plot_65 <- plot_grid(pred_contacts_65, pred_mask_65, pred_vaccines_65, ncol = 1,labels = c("A", "B", "C"))

combined_plot_65 <- plot_grid(
  ggdraw() + draw_label("", size = 16, hjust = 0, x = 0.01),
  combined_plot_65,
  ncol = 1, 
  rel_heights = c(0.4, 10) 
)

combined_plot_65

ggsave("combined_contrasts_65_plus.png", plot = combined_plot_65, path = "../out/plots", width = 12, height = 12, units = "in")

rm(combined_plot_65, pred_contacts_65, pred_mask_65, pred_vaccines_65)
```


# =============================================================================
# 6. WAVE-LEVEL DIFFERENCES
# =============================================================================

Let's plot differences by wave now:
```{r}
average_by_wave_dems <- D_df %>%
  group_by(data_collected_dates) %>%
  summarise(
    average_num_cc = weighted.mean(num_cc, w = weight_party_raked, na.rm = TRUE),
    average_mask_usage = weighted.mean(Norm_Masks_Used, w = weight_party_raked, na.rm = TRUE),
    average_safety_index = weighted.mean(Non_Weighted_Safety_Index, w = weight_party_raked, na.rm = TRUE),
    average_vaccinated = weighted.mean(Vaccinated, w = weight_party_raked, na.rm = TRUE)
  )

average_by_wave_repubs <- R_df %>%
  group_by(data_collected_dates) %>%
  summarise(
    average_num_cc = weighted.mean(num_cc, w = weight_party_raked, na.rm = TRUE),
    average_mask_usage = weighted.mean(Norm_Masks_Used, w = weight_party_raked, na.rm = TRUE),
    average_safety_index = weighted.mean(Non_Weighted_Safety_Index, w = weight_party_raked, na.rm = TRUE),
    average_vaccinated = weighted.mean(Vaccinated, w = weight_party_raked, na.rm = TRUE)
  )

#identifying the parties
average_by_wave_dems$group <- 'Democrats'
average_by_wave_repubs$group <- 'Republicans'

average_by_wave <- rbind(average_by_wave_dems, average_by_wave_repubs)

# Calculate the differences
average_diff_by_wave <- average_by_wave_repubs %>%
  inner_join(average_by_wave_dems, by = "data_collected_dates", suffix = c("_repubs", "_dems")) %>%
  mutate(
    average_diff_contacts = average_num_cc_repubs - average_num_cc_dems,
    average_diff_masks = average_mask_usage_repubs - average_mask_usage_dems,
    average_diff_safety = average_safety_index_repubs - average_safety_index_dems,
    average_diff_vaccines = average_vaccinated_repubs - average_vaccinated_dems
  ) %>%
  select(data_collected_dates, starts_with("average_diff"))

write_csv(average_by_wave, "../data/party_average_by_wave.csv")
write_csv(average_diff_by_wave, "../data/party_average_by_wave_differences.csv")

average_by_wave
average_diff_by_wave

rvd_masks_diff_by_wave <- ggplot(average_diff_by_wave, aes(x = factor(data_collected_dates), y = average_diff_masks)) +
  geom_bar(stat = "identity", fill = "#c50000", color = "black", alpha = 0.7) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "black") +
  labs(title = "Difference in Percent of Contacts with a Mask (Republicans - Democrats) \nOver Time",
       x = "Data Collection Period",
       y = "") +
  theme_cowplot() +
  theme(
    plot.title = element_text(size = 22),
    panel.grid.minor.y = element_blank(),
    panel.grid.major.x = element_line(color = "grey70", size = 0.5),
    panel.grid.major.y = element_line(color = "grey90", size = 0.5),
    axis.title.x = element_blank(),
    axis.text.x = element_text(size = 14, angle = 45, hjust = 1, vjust = 1),
    axis.text.y = element_text(size = 22),
    axis.title.y = element_text(size = 16),
    legend.title = element_text(size = 12),
    legend.text = element_text(size = 16),
    plot.margin = margin(10, 10, 10, 10)
  ) +
  guides(shape = guide_legend(override.aes = list(size = 6)),
         color = "none", 
         size = "none")

ggsave("rvd_difference_masks_waves.png", plot = rvd_masks_diff_by_wave, path = "../out/plots", width = 12, height = 8, units = "in")

rvd_contacts_diff_by_wave <-  ggplot(average_diff_by_wave, aes(x = factor(data_collected_dates), y = average_diff_contacts)) +
  geom_bar(stat = "identity", fill = "#c50000", color = "black", alpha = 0.7) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "black") +
  labs(title = "Difference in Average Contacts (Republicans - Democrats) Over Time",
       x = "Data Collection Period",
       y = "") +
  theme_cowplot() +
  theme(
    plot.title = element_text(size = 22),
    panel.grid.minor.y = element_blank(),
    panel.grid.major.x = element_line(color = "grey70", size = 0.5),
    panel.grid.major.y = element_line(color = "grey90", size = 0.5),
    axis.title.x = element_blank(),
    axis.text.x = element_text(size = 14, angle = 45, hjust = 1, vjust = 1),
    axis.text.y = element_text(size = 22),
    axis.title.y = element_text(size = 16),
    legend.title = element_text(size = 12),
    legend.text = element_text(size = 16),
    plot.margin = margin(10, 10, 10, 10)
  ) +
  guides(shape = guide_legend(override.aes = list(size = 6)),
         color = "none", 
         size = "none")

ggsave("rvd_difference_contacts_waves.png", plot = rvd_contacts_diff_by_wave, path = "../out/plots", width = 12, height = 8, units = "in")

rm(average_by_wave, average_by_wave_dems, average_by_wave_repubs, average_diff_by_wave, rvd_masks_contacts_by_wave, rvd_masks_diff_by_wave, bi_college, bi_concern, bi_independent, contr_college, contr_independent, contr_republican, df_college_contrasts, df_independent_contrasts, df_republican_contrasts, emm_college, emm_concern, emm_independent, emm_republican, rvd_contacts_diff_by_wave, bi_republican)
```

