---
title: "03-Waves_Pooled_Behavior_Regressions"
author: "Christopher Soria"
date: "2024-01-30"
output: html_document
---

# =============================================================================
# TABLE OF CONTENTS
# -----------------------------------------------------------------------------
# 1. SETUP AND CONFIGURATION
# 2. CONTACTS REGRESSIONS
# 3. MASK USAGE REGRESSIONS
# 4. VACCINATION REGRESSIONS
# 5. COMBINED PLOTS
# 6. REGRESSION TABLES
# =============================================================================

# =============================================================================
# 1. SETUP AND CONFIGURATION
# =============================================================================

```{r setup, message=FALSE, warning=FALSE}
library(stargazer)
library(tidyverse)
library(emmeans)
library(cowplot)
library(broom)

# Source helper functions
source("plot_helper_functions/regression_helpers.R")
source("plot_helper_functions/plot_utils.R")

emm_options(rg.limit = 775000)
dodge_width <- .7

# Load Data
if (!exists("bics_zip_features")) {
  if (file.exists("../data/BICS_ZIP_Features.csv")) {
    bics_zip_features <- read_csv("../data/BICS_ZIP_Features.csv")
  } else {
    stop("Data file ../data/BICS_ZIP_Features.csv not found. Please run 00-Prep_BICS_Data.Rmd first.")
  }
}

# Calculate averages for emmeans
average_hhsize <- weighted.mean(bics_zip_features$resp_hhsize, w = bics_zip_features$weight_party_raked, na.rm = TRUE)
average_mort <- weighted.mean(bics_zip_features$log_prev_week_mort_rate, w = bics_zip_features$weight_party_raked, na.rm = TRUE)
average_CD_dem <- weighted.mean(bics_zip_features$CD_PERCENT_DEMOCRAT, w = bics_zip_features$weight_party_raked, na.rm = TRUE)
```

# =============================================================================
# 2. CONTACTS REGRESSIONS
# =============================================================================


##Contacts
```{r contacts}
# Run nested models and generate emmeans
contacts_results <- run_nested_models(
  data = bics_zip_features,
  outcome_var = "num_cc",
  average_hhsize = average_hhsize,
  average_mort = average_mort,
  average_CD_dem = average_CD_dem
)

print(summary(contacts_results$model1))
print(summary(contacts_results$model2))
print(summary(contacts_results$model3))

# Plot all three models
rvd_contacts_plot <- plot_emmeans_models(
  contacts_results$all_emms,
  title = "predicted number of daily contacts"
)

rvd_contacts_plot

# Plot model 3 only
rvd_contacts_plot_emm_3 <- plot_single_model_emmeans(
  contacts_results$emm3,
  title = "Number of daily contacts",
  x_label = "Daily Contacts"
)

rvd_contacts_plot_emm_3

# Save plots
ggsave("rvd_contacts_pooled.png", plot = rvd_contacts_plot, path = "../plots/Contacts", width = 12, height = 8, units = "in")
ggsave("rvd_contacts_pooled_only_two_models.png", plot = rvd_contacts_plot_emm_3, path = "../plots/Contacts", width = 12, height = 8, units = "in")
```

# =============================================================================
# 3. MASK USAGE REGRESSIONS
# =============================================================================

##Mask Usage
```{r masks}
# Run nested models and generate emmeans
masks_results <- run_nested_models(
  data = bics_zip_features,
  outcome_var = "Norm_Masks_Used",
  average_hhsize = average_hhsize,
  average_mort = average_mort,
  average_CD_dem = average_CD_dem
)

print(summary(masks_results$model1))
print(summary(masks_results$model2))
print(summary(masks_results$model3))

# Plot all three models
rvd_masks_plot <- plot_emmeans_models(
  masks_results$all_emms,
  title = "Percentage of contacts with a mask"
)

rvd_masks_plot

# Plot model 3 only
rvd_masks_plot_emm_3 <- plot_single_model_emmeans(
  masks_results$emm3,
  title = "Percent contacts with a mask",
  x_label = "Mask Usage"
)

rvd_masks_plot_emm_3

# Save plots
ggsave("rvd_masks_pooled.png", plot = rvd_masks_plot, path = "../plots/Masks", width = 12, height = 8, units = "in")
ggsave("rvd_masks_pooled_only_two_models.png", plot = rvd_masks_plot_emm_3, path = "../plots/Masks", width = 12, height = 8, units = "in")
```

# =============================================================================
# 4. VACCINATION REGRESSIONS
# =============================================================================

##Vaccines
```{r vaccines}
# Run nested models and generate emmeans
vaccination_results <- run_nested_models(
  data = bics_zip_features,
  outcome_var = "Vaccinated",
  average_hhsize = average_hhsize,
  average_mort = average_mort,
  average_CD_dem = average_CD_dem,
  include_wave = FALSE
)

print(summary(vaccination_results$model1))
print(summary(vaccination_results$model2))
print(summary(vaccination_results$model3))

# Plot all three models
rvd_vaccination_plot <- plot_emmeans_models(
  vaccination_results$all_emms,
  title = "predicted probability of vaccination"
)

rvd_vaccination_plot

# Save plot
ggsave("rvd_vaccination_pooled.png", plot = rvd_vaccination_plot, path = "../plots/Vaccines", width = 12, height = 8, units = "in")
```

# vaccines only 2 models
```{r vaccines_model3}
# Plot model 3 only
rvd_vaccination_plot_12 <- plot_single_model_emmeans(
  vaccination_results$emm3,
  title = "Probability of vaccination",
  x_label = "Vaccination"
)

rvd_vaccination_plot_12

ggsave("rvd_vaccination_pooled_12.png",
  plot = rvd_vaccination_plot_12,
  path = "../plots/Vaccines",
  width = 12, height = 8, units = "in"
)
```



# =============================================================================
# 5. COMBINED PLOTS
# =============================================================================

##Combined Plot
```{r}
legend <- get_legend(
  rvd_contacts_plot +
    theme(
      legend.position = "bottom",
      legend.text = element_text(size = 20),
      legend.title = element_text(size = 20),
      legend.margin = margin(t = 10, r = 10, b = 10, l = 10)
    )
)

combined_plot <- plot_grid(
  plot_grid(rvd_contacts_plot + theme(legend.position = "none"),
    rvd_masks_plot + theme(legend.position = "none"),
    rvd_vaccination_plot + theme(legend.position = "none"),
    ncol = 1, labels = c("A", "B", "C")
  ),
  legend,
  ncol = 1,
  rel_heights = c(1, 0.1) # Adjust the relative height of the legend
)

combined_plot <- plot_grid(
  combined_plot,
  ncol = 1,
  rel_heights = c(0.4, 10)
)

ggsave("combined_national_pooled.png", plot = combined_plot, path = "../out/plots", width = 12, height = 14, units = "in")

combined_plot

rm(combined_plot, rvd_contacts_plot, rvd_masks_plot, rvd_vaccination_plot, legend)
```
# Combined plot with only model 3
```{r}
legend <- get_legend_35(
  rvd_contacts_plot_emm_3 +
    theme(
      legend.position = "bottom",
      legend.text = element_text(size = 20),
      legend.title = element_text(size = 20),
      legend.margin = margin(t = 10, r = 10, b = 10, l = 10)
    )
)

combined_plot <- plot_grid(
  plot_grid(rvd_contacts_plot_emm_3 + theme(legend.position = "none"),
    rvd_masks_plot_emm_3 + theme(legend.position = "none"),
    rvd_vaccination_plot_12 + theme(legend.position = "none"),
    ncol = 3, labels = c("A", "B", "C")
  ),
  legend,
  ncol = 1,
  rel_heights = c(1, 0.1)
)

combined_plot <- plot_grid(
  combined_plot,
  ncol = 1,
  rel_heights = c(0.4, 10)
)

ggsave("combined_national_pooled_two_mdoels.png", plot = combined_plot, path = "../out/plots", width = 16, height = 6, units = "in")

combined_plot

# rm(combined_plot, rvd_contacts_plot_emm_3, rvd_masks_plot_emm_3, rvd_vaccination_plot_12, legend)
```



# =============================================================================
# 6. REGRESSION TABLES
# =============================================================================

#Combined Tables
```{r}
custom_labels <- c(
  "political_partyIndependent" = "Independent Party (ref: Democrat)",
  "political_partyRepublican" = "Republican Party (ref: Democrat)",
  "political_partyPrefer not to answer" = "Prefer Not to Answer (ref: Democrat)",
  "wave2" = "Wave 2 (ref: Wave 1)",
  "wave3" = "Wave 3 (ref: Wave 1)",
  "wave4" = "Wave 4  (ref: Wave 1)",
  "wave5" = "Wave 5  (ref: Wave 1)",
  "wave6" = "Wave 6  (ref: Wave 1)",
  "interview_day_weekendWeekend" = "Weekend Interview",
  "age_group[30,45)" = "Age Group: [30,45) (ref: [18,30)",
  "age_group[45,55)" = "Age Group: [45,55) (ref: [18,30)",
  "age_group[55,65)" = "Age Group: [55,65) (ref: [18,30)",
  "age_group[65,100]" = "Age Group: [65,100] (ref: [18,30)",
  "r_raceBlack" = "Race: Black (ref: White)",
  "r_raceAsian" = "Race: Asian (ref: White)",
  "r_raceOther / Mixed" = "Race: Other/Mixed (ref: White)",
  "resp_hispanic.L" = "Hispanic",
  "r_male1" = "Male",
  "resp_educHigh school graduate" = "High School Graduate (ref: College Grad)",
  "resp_educLess than high school" = "Less than High School (ref: College Grad)",
  "r_workingWorking" = "Employed",
  "resp_hhsize" = "Household Size",
  "r_metro1" = "Metro Area",
  "CD_PERCENT_DEMOCRAT" = "CD Percent Democrat",
  "log_prev_week_mort_rate" = "Log Previous Week Incidence Rate",
  "county_mask_mandateLess Strict" = "Less Strict Mask Mandate (ref: Strict Mandate)",
  "county_mask_mandateUnspecified or Unclear Mandate" = "Unspecified Mask Mandate (ref: Strict Mandate)",
  "county_mask_mandateNo Mandate" = "No Mask Mandate (ref: Strict Mandate)",
  "Constant" = "Constant"
)

custom_labels_vaccines <- c(
  "political_partyIndependent" = "Independent Party (ref: Democrat)",
  "political_partyRepublican" = "Republican Party (ref: Democrat)",
  "political_partyPrefer not to answer" = "Prefer Not to Answer (ref: Democrat)",
  "interview_day_weekendWeekend" = "Weekend Interview",
  "age_group[30,45)" = "Age Group: [30,45) (ref: [18,30)",
  "age_group[45,55)" = "Age Group: [45,55) (ref: [18,30)",
  "age_group[55,65)" = "Age Group: [55,65) (ref: [18,30)",
  "age_group[65,100]" = "Age Group: [65,100] (ref: [18,30)",
  "r_raceBlack" = "Race: Black (ref: White)",
  "r_raceAsian" = "Race: Asian (ref: White)",
  "r_raceOther / Mixed" = "Race: Other/Mixed (ref: White)",
  "resp_hispanic.L" = "Hispanic",
  "r_male1" = "Male",
  "resp_educHigh school graduate" = "High School Graduate (ref: College Grad)",
  "resp_educLess than high school" = "Less than High School (ref: College Grad)",
  "r_workingWorking" = "Employed",
  "resp_hhsize" = "Household Size",
  "r_metro1" = "Metro Area",
  "CD_PERCENT_DEMOCRAT" = "CD Percent Democrat",
  "log_prev_week_mort_rate" = "Log Previous Week Incidence Rate",
  "county_mask_mandateLess Strict" = "Less Strict Mask Mandate (ref: Strict Mandate)",
  "county_mask_mandateUnspecified or Unclear Mandate" = "Unspecified Mask Mandate (ref: Strict Mandate)",
  "county_mask_mandateNo Mandate" = "No Mask Mandate (ref: Strict Mandate)",
  "Constant" = "Constant"
)

reference_lines <- list(
  c("Reference Categories", ""),
  c("Political Party", "Democrat"),
  c("Male", "Non-Male"),
  c("Wave", "Wave 2"),
  c("Weekend Interview", "Weekday Interview"),
  c("Education", "College Graduate"),
  c("Employed", "Not Employed"),
  c("Metro Area", "Non-Metro Area"),
  c("County Mask Mandate", "Strict"),
  c("\\hline", "") # Add a horizontal line after reference categories
)

reference_lines_vaccines <- list(
  c("reference categories", ""),
  c("political party", "democrat"),
  c("male", "non-male"),
  c("weekend interview", "weekday interview"),
  c("education", "college graduate"),
  c("employed", "not employed"),
  c("metro area", "non-metro area"),
  c("county mask mandate", "strict"),
  c("\\hline", "")
)

stargazer(contacts_results$model1, contacts_results$model2, contacts_results$model3,
  type = "latex",
  title = "Association between Political Party and the Number of Reported Contacts",
  align = TRUE,
  no.space = TRUE,
  covariate.labels = custom_labels,
  omit = c("STATE", "city"),
  column.labels = c("Model 1", "Model 2", "Model 3"),
  dep.var.labels = "Total Contacts",
  notes = "\\footnotesize",
  out = "../plots/tables/contacts_regression_results.tex"
)


stargazer(masks_results$model1, masks_results$model2, masks_results$model3,
  type = "latex",
  title = "Association between Political Party and Reported Percentage of Contacts with a Mask",
  align = TRUE,
  no.space = TRUE,
  covariate.labels = custom_labels,
  omit = c("STATE", "city"),
  dep.var.labels = "Percentage of Contacts with a Mask",
  notes = "\\footnotesize",
  out = "../plots/tables/masks_regression_results.tex"
)


suppressWarnings(
  stargazer(vaccination_results$model1, vaccination_results$model2, vaccination_results$model3,
    type = "latex",
    title = "Association between Political Party and Reported Vaccination",
    align = TRUE,
    no.space = TRUE,
    covariate.labels = custom_labels_vaccines,
    omit = c("STATE", "city"),
    column.labels = c("Model 1", "Model 2", "Model 3"),
    dep.var.labels = "Probability of Vaccination",
    notes = "\\footnotesize",
    out = "../plots/tables/vaccines_regression_results.tex"
  )
)


# Cleanup
rm(contacts_results, masks_results, vaccination_results)
rm(rvd_contacts_plot, rvd_masks_plot, rvd_vaccination_plot)
rm(rvd_contacts_plot_emm_3, rvd_masks_plot_emm_3, rvd_vaccination_plot_12)
```
```{r cleanup}
rm(average_CD_dem, average_hhsize, average_mort, dodge_width, custom_labels, custom_labels_vaccines, reference_lines)
```

