---
title: "05-incidence_mortality_response"
author: "Christopher Soria"
date: "2024-10-16"
output: html_document
---

# =============================================================================
# TABLE OF CONTENTS
# -----------------------------------------------------------------------------
# 1. SETUP AND CONFIGURATION
# 2. CONTACTS RESPONSE TO INCIDENCE (OVERALL)
# 3. CONTACTS RESPONSE TO MORTALITY (OVERALL)
# 4. COMBINED: CONTACTS OVERALL
# 5. CONTACTS RESPONSE TO INCIDENCE (BY PARTY)
# 6. CONTACTS RESPONSE TO MORTALITY (BY PARTY)
# 7. COMBINED: CONTACTS BY PARTY
# 8. CONTACTS BY WAVE AND MORTALITY
# 9. MASKS RESPONSE TO INCIDENCE (BY PARTY)
# 10. MASKS RESPONSE TO MORTALITY (BY PARTY)
# 11. COMBINED: MASKS BY PARTY
# 12. CLEANUP
# =============================================================================

# =============================================================================
# 1. SETUP AND CONFIGURATION
# =============================================================================

```{r setup, message=FALSE, warning=FALSE}
library(stargazer)
library(tidyverse)
library(emmeans)
library(cowplot)
library(broom)

emm_options(rg.limit = 775000)
dodge_width <- .7

# Load Data
if (!exists("bics_zip_features")) {
  if (file.exists("../data/BICS_ZIP_Features.csv")) {
    bics_zip_features <- read_csv("../data/BICS_ZIP_Features.csv")
  } else {
    stop("Data file ../data/BICS_ZIP_Features.csv not found. Please run 00-Prep_BICS_Data.Rmd first.")
  }
}

# Calculate weighted averages for emmeans
average_hhsize <- weighted.mean(bics_zip_features$resp_hhsize, w = bics_zip_features$weight_party_raked, na.rm = TRUE)
average_CD_dem <- weighted.mean(bics_zip_features$CD_PERCENT_DEMOCRAT, w = bics_zip_features$weight_party_raked, na.rm = TRUE)
```

Below, I want to find out who is more responsive to incidence and mortality rates (Democrats or Republicans).

# =============================================================================
# 2. CONTACTS RESPONSE TO INCIDENCE (OVERALL)
# =============================================================================
```{r}
contacts_interaction_model1 <- lm(num_cc ~ log_prev_week_inc_rate + wave + interview_day_weekend + city,
                          data = bics_zip_features,
                          weights = weight_party_raked)

print(summary(contacts_interaction_model1))

contacts_interaction_model2 <- lm(num_cc ~ log_prev_week_inc_rate + wave + interview_day_weekend + city + age_group + r_race + resp_hispanic + r_male + educ_group + r_working + resp_hhsize, 
                                   data = bics_zip_features, 
                                   weights = weight_party_raked)


print(summary(contacts_interaction_model2))


contacts_interaction_model3 <- lm(num_cc ~ log_prev_week_inc_rate + county_mask_mandate + r_metro + male + educ_group + age_group + resp_hhsize + r_race + resp_hispanic + r_working + city + wave + interview_day_weekend + CD_PERCENT_DEMOCRAT + STATE, 
                                   data = bics_zip_features, 
                                   weights = weight_party_raked)

print(summary(contacts_interaction_model3))


log_prev_week_inc_rate_values <- seq(from = min(bics_zip_features$log_prev_week_inc_rate, na.rm = TRUE),
                                     to = max(bics_zip_features$log_prev_week_inc_rate, na.rm = TRUE),
                                     length.out = 100)

emm_contacts1 <- emmeans(
  contacts_interaction_model1,
  specs = "log_prev_week_inc_rate",
  at = list(log_prev_week_inc_rate = log_prev_week_inc_rate_values,
            wave = "2",
            interview_day_weekend = "Weekday",
            city = "National")
)


average_yob <- weighted.mean(bics_zip_features$age_group, w = bics_zip_features$weight_party_raked, na.rm = TRUE)
average_hhsize <- weighted.mean(bics_zip_features$resp_hhsize, w = bics_zip_features$weight_party_raked, na.rm = TRUE)

emm_contacts2 <- emmeans(
  contacts_interaction_model2,
  specs = "log_prev_week_inc_rate",
  at = list(log_prev_week_inc_rate = log_prev_week_inc_rate_values,
            wave = "2",
            interview_day_weekend = "Weekday",
            city = "National",
            age_group = "[18,30)",
            r_race = "White",
            resp_hispanic = "No",
            r_male = "0",
            r_working = "Working",
            resp_hhsize = average_hhsize,
            educ_group = "College graduate and above")
)


average_CD_dem <- weighted.mean(bics_zip_features$CD_PERCENT_DEMOCRAT, w = bics_zip_features$weight_party_raked, na.rm = TRUE)

emm_contacts3 <- emmeans(
  contacts_interaction_model3,
  specs = "log_prev_week_inc_rate",
  at = list(log_prev_week_inc_rate = log_prev_week_inc_rate_values,
    wave = "2",
            interview_day_weekend = "Weekday",
            city = "National",
            age_group = "[18,30)",
            r_race = "White",
            resp_hispanic = "No",
            r_male = "0",
            r_working = "Working",
            resp_hhsize = average_hhsize,
            educ_group = "College graduate and above",
            county_mask_mandate = "Strict",
            r_metro = "1",
            CD_PERCENT_DEMOCRAT = average_CD_dem,
            STATE = "California")
)


df_emm_contacts1 <- as.data.frame(summary(emm_contacts1))
df_emm_contacts2 <- as.data.frame(summary(emm_contacts2))
df_emm_contacts3 <- as.data.frame(summary(emm_contacts3))

df_emm_contacts1$model <- "1"
df_emm_contacts2$model <- "2"
df_emm_contacts3$model <- "3"

all_emms <- rbind(df_emm_contacts1, df_emm_contacts2, df_emm_contacts3)

contacts_interaction_plot <- ggplot(all_emms, aes(x = log_prev_week_inc_rate, y = emmean)) +
  geom_line() +
  geom_ribbon(aes(ymin = emmean - SE, ymax = emmean + SE), alpha = 0.5, linetype = "blank") +
  facet_wrap(~ model) +
  scale_fill_manual(values = c("Independent" = "#6db753", "Republican" = "#c50000", "Democrat" = "#326ba4")) +
  labs(
    title = "Impact on Daily Contacts",
    x = "Log Prev Week Inc Rate",
    y = "Estimated Marginal Means",
    color = "Political Party",
    fill = "Political Party"
  ) +
  theme_cowplot() +
  theme(
    strip.text = element_blank(),
    plot.title = element_text(size = 22),
    axis.title.x = element_blank(),
    axis.text.x = element_text(size = 15),
    axis.text.y = element_text(size = 22),
    axis.title.y = element_blank(),
    legend.title = element_text(size = 12),
    legend.text = element_text(size = 16),
    panel.grid.major.y = element_line(color = "grey", size = 0.5)
  )

contacts_interaction_plot

mean_log_prev_week_inc_rate <- mean(log_prev_week_inc_rate_values, na.rm = TRUE)
log_prev_week_inc_rate_values_10pct <- mean_log_prev_week_inc_rate * 1.10

emm_contacts1 <- emmeans(
  contacts_interaction_model1,
  specs = "log_prev_week_inc_rate",
  at = list(log_prev_week_inc_rate = mean_log_prev_week_inc_rate,
            wave = "2",
            interview_day_weekend = "Weekday",
            city = "National")
)

emm_contacts1_10pct <- emmeans(
  contacts_interaction_model1,
  specs = "log_prev_week_inc_rate",
  at = list(log_prev_week_inc_rate = log_prev_week_inc_rate_values_10pct,
            wave = "2",
            interview_day_weekend = "Weekday",
            city = "National")
)

combined_emm <- rbind(mean_value = emm_contacts1, increased_value = emm_contacts1_10pct)

contrast_results <- contrast(combined_emm, method = "revpairwise")

contrast_df <- as.data.frame(contrast_results)

contacts_interaction_diff_plot <- ggplot(contrast_df, aes(x = contrast, y = estimate, fill = contrast)) +
  geom_bar(stat = "identity", alpha = 0.7, fill = "black", width = 0.5) +
  geom_errorbar(aes(ymin = estimate - SE, ymax = estimate + SE), width = 0.2, color = "black") +
  labs(
    title = "Difference in Contacts Associated with a 10% Increase in Previous Week Incidence",
    x = "",
    y = "Difference in Estimated Marginal Means"
  ) +
  #ylim(-1.1, 0.11) +
  theme_cowplot() +
  theme(
    strip.text = element_blank(),
    plot.title = element_text(size=22),
    axis.title.x = element_blank(),
    axis.title.y = element_blank(),
    axis.text.x = element_blank(),
    axis.text.y = element_text(size = 15),
    legend.position="none", 
    panel.grid.major.y=element_line(color="grey", size=0.5)
  )

print(contacts_interaction_diff_plot)
```

# =============================================================================
# 3. CONTACTS RESPONSE TO MORTALITY (OVERALL)
# =============================================================================
```{r}
contacts_interaction_model1 <- lm(num_cc ~ log_prev_week_mort_rate + wave + interview_day_weekend + city,
                          data = bics_zip_features,
                          weights = weight_party_raked)

print(summary(contacts_interaction_model1))

contacts_interaction_model2 <- lm(num_cc ~ log_prev_week_mort_rate + wave + interview_day_weekend + city + age_group + r_race + resp_hispanic + r_male + educ_group + r_working + resp_hhsize, 
                                   data = bics_zip_features, 
                                   weights = weight_party_raked)


print(summary(contacts_interaction_model2))


contacts_interaction_model3 <- lm(num_cc ~ log_prev_week_mort_rate + log_prev_week_mort_rate + county_mask_mandate + r_metro + male + educ_group + age_group + resp_hhsize + r_race + resp_hispanic + r_working + city + wave + interview_day_weekend + CD_PERCENT_DEMOCRAT + STATE, 
                                   data = bics_zip_features, 
                                   weights = weight_party_raked)

print(summary(contacts_interaction_model3))


log_prev_week_mort_rate_values <- seq(from = min(bics_zip_features$log_prev_week_mort_rate, na.rm = TRUE),
                                     to = max(bics_zip_features$log_prev_week_mort_rate, na.rm = TRUE),
                                     length.out = 100)

emm_contacts1 <- emmeans(
  contacts_interaction_model1,
  specs = "log_prev_week_mort_rate",
  at = list(log_prev_week_mort_rate = log_prev_week_mort_rate_values,
            wave = "2",
            interview_day_weekend = "Weekday",
            city = "National")
)


average_yob <- weighted.mean(bics_zip_features$age_group, w = bics_zip_features$weight_party_raked, na.rm = TRUE)
average_hhsize <- weighted.mean(bics_zip_features$resp_hhsize, w = bics_zip_features$weight_party_raked, na.rm = TRUE)

emm_contacts2 <- emmeans(
  contacts_interaction_model2,
  specs = "log_prev_week_mort_rate",
  at = list(log_prev_week_mort_rate = log_prev_week_mort_rate_values,
            wave = "2",
            interview_day_weekend = "Weekday",
            city = "National",
            age_group = "[18,30)",
            r_race = "White",
            resp_hispanic = "Yes",
            r_male = "0",
            r_working = "Working",
            resp_hhsize = average_hhsize,
            educ_group = "College graduate and above")
)


average_CD_dem <- weighted.mean(bics_zip_features$CD_PERCENT_DEMOCRAT, w = bics_zip_features$weight_party_raked, na.rm = TRUE)

emm_contacts3 <- emmeans(
  contacts_interaction_model3,
  specs = "log_prev_week_mort_rate",
  at = list(log_prev_week_mort_rate = log_prev_week_mort_rate_values,
    wave = "2",
            interview_day_weekend = "Weekday",
            city = "National",
            age_group = "[18,30)",
            r_race = "White",
            resp_hispanic = "Yes",
            r_male = "0",
            r_working = "Working",
            resp_hhsize = average_hhsize,
            educ_group = "College graduate and above",
            county_mask_mandate = "Strict",
            r_metro = "1",
            CD_PERCENT_DEMOCRAT = average_CD_dem,
            STATE = "California")
)


df_emm_contacts1 <- as.data.frame(summary(emm_contacts1))
df_emm_contacts2 <- as.data.frame(summary(emm_contacts2))
df_emm_contacts3 <- as.data.frame(summary(emm_contacts3))

df_emm_contacts1$model <- "1"
df_emm_contacts2$model <- "2"
df_emm_contacts3$model <- "3"

all_emms <- rbind(df_emm_contacts1, df_emm_contacts2, df_emm_contacts3)

contacts_interaction_plot <- ggplot(all_emms, aes(x = log_prev_week_mort_rate, y = emmean)) +
  geom_line() +
  geom_ribbon(aes(ymin = emmean - SE, ymax = emmean + SE), alpha = 0.5, linetype = "blank") +
  facet_wrap(~ model) +
  scale_fill_manual(values = c("Independent" = "#6db753", "Republican" = "#c50000", "Democrat" = "#326ba4")) +
  labs(
    title = "Impact on Daily Contacts",
    x = "Log Prev Week Mort Rate",
    y = "Estimated Marginal Means",
    color = "Political Party",
    fill = "Political Party"
  ) +
  theme_cowplot() +
  theme(
    strip.text = element_blank(),
    plot.title = element_text(size = 22),
    axis.title.x = element_blank(),
    axis.text.x = element_text(size = 15),
    axis.text.y = element_text(size = 22),
    axis.title.y = element_blank(),
    legend.title = element_text(size = 12),
    legend.text = element_text(size = 16),
    panel.grid.major.y = element_line(color = "grey", size = 0.5)
  )

contacts_interaction_plot

mean_log_prev_week_mort_rate <- mean(log_prev_week_mort_rate_values, na.rm = TRUE)
log_prev_week_mort_rate_values_10pct <- mean_log_prev_week_mort_rate * 1.10

emm_contacts1 <- emmeans(
  contacts_interaction_model1,
  specs = "log_prev_week_mort_rate",
  at = list(log_prev_week_mort_rate = mean_log_prev_week_mort_rate,
            wave = "2",
            interview_day_weekend = "Weekday",
            city = "National")
)

emm_contacts1_10pct <- emmeans(
  contacts_interaction_model1,
  specs = "log_prev_week_mort_rate",
  at = list(log_prev_week_mort_rate = log_prev_week_mort_rate_values_10pct,
            wave = "2",
            interview_day_weekend = "Weekday",
            city = "National")
)

combined_emm <- rbind(mean_value = emm_contacts1, increased_value = emm_contacts1_10pct)

contrast_results <- contrast(combined_emm, method = "revpairwise")

contrast_df <- as.data.frame(contrast_results)

contacts_interaction_diff_plot_mort <- ggplot(contrast_df, aes(x = contrast, y = estimate, fill = contrast)) +
  geom_bar(stat = "identity", alpha = 0.7, fill = "black", width = 0.5) +
  geom_errorbar(aes(ymin = estimate - SE, ymax = estimate + SE), width = 0.2, color = "black") +
  labs(
    title = "Difference in Contacts Associated with a 10% Increase in Previous Week Mortality",
    x = "",
    y = "Difference in Estimated Contacts"
  ) +
  #ylim(-1.1, 0.11) +
  theme_cowplot() +
  theme(
    strip.text = element_blank(),
    plot.title = element_text(size=22),
    axis.title.x = element_blank(),
    axis.title.y = element_blank(),
    axis.text.x = element_blank(),
    axis.text.y = element_text(size = 15),
    legend.position="none", 
    panel.grid.major.y=element_line(color="grey", size=0.5)
  )

print(contacts_interaction_diff_plot_mort)
```

# =============================================================================
# 4. COMBINED: CONTACTS OVERALL
# =============================================================================
```{r}
contacts_interaction_diff_plot_mort <- contacts_interaction_diff_plot_mort + 
  theme(legend.position = "none")

combined_plot <- plot_grid(
  plot_grid(contacts_interaction_diff_plot + theme(legend.position = "none"),
            contacts_interaction_diff_plot_mort,
            ncol = 1, labels = c("A", "B")),
  ncol = 1
)

combined_plot <- plot_grid(
  ggdraw() + draw_label("Figure", size = 16, hjust = 0, x = 0.01),
  combined_plot,
  ncol = 1,
  rel_heights = c(0.4, 10)
)

print(combined_plot)

ggsave("combined_national_pooled_mortalty_and_incidence_response_contacts_overall.png", plot = combined_plot, path = "../out/plots", width = 14, height = 16, units = "in")
```

# =============================================================================
# 5. CONTACTS RESPONSE TO INCIDENCE (BY PARTY)
# =============================================================================
```{r}
rvd_contacts_interaction_model1 <- lm(num_cc ~ political_party*log_prev_week_inc_rate + wave + interview_day_weekend + city,
                          data = bics_zip_features,
                          weights = weight_party_raked)

print(summary(rvd_contacts_interaction_model1))

rvd_contacts_interaction_model2 <- lm(num_cc ~ political_party*log_prev_week_inc_rate + wave + interview_day_weekend + city + age_group + r_race + resp_hispanic + r_male + educ_group + r_working + resp_hhsize, 
                                   data = bics_zip_features, 
                                   weights = weight_party_raked)


print(summary(rvd_contacts_interaction_model2))


rvd_contacts_interaction_model3 <- lm(num_cc ~ political_party*log_prev_week_inc_rate + log_prev_week_inc_rate + county_mask_mandate + r_metro + male + educ_group + age_group + resp_hhsize + r_race + resp_hispanic + r_working + city + wave + interview_day_weekend + CD_PERCENT_DEMOCRAT + STATE, 
                                   data = bics_zip_features, 
                                   weights = weight_party_raked)

print(summary(rvd_contacts_interaction_model3))


log_prev_week_inc_rate_values <- seq(from = min(bics_zip_features$log_prev_week_inc_rate, na.rm = TRUE),
                                     to = max(bics_zip_features$log_prev_week_inc_rate, na.rm = TRUE),
                                     length.out = 100)

emm_contacts1 <- emmeans(
  rvd_contacts_interaction_model1,
  specs = "political_party",
  by = "log_prev_week_inc_rate",
  at = list(log_prev_week_inc_rate = log_prev_week_inc_rate_values,
            wave = "2",
            interview_day_weekend = "Weekday",
            city = "National")
)


average_yob <- weighted.mean(bics_zip_features$age_group, w = bics_zip_features$weight_party_raked, na.rm = TRUE)
average_hhsize <- weighted.mean(bics_zip_features$resp_hhsize, w = bics_zip_features$weight_party_raked, na.rm = TRUE)

emm_contacts2 <- emmeans(
  rvd_contacts_interaction_model2,
  specs = "political_party",
  by = "log_prev_week_inc_rate",
  at = list(log_prev_week_inc_rate = log_prev_week_inc_rate_values,
            wave = "2",
            interview_day_weekend = "Weekday",
            city = "National",
            age_group = "[18,30)",
            r_race = "White",
            resp_hispanic = "No",
            r_male = "0",
            r_working = "Working",
            resp_hhsize = average_hhsize,
            educ_group = "College graduate and above")
)


average_CD_dem <- weighted.mean(bics_zip_features$CD_PERCENT_DEMOCRAT, w = bics_zip_features$weight_party_raked, na.rm = TRUE)

emm_contacts3 <- emmeans(
  rvd_contacts_interaction_model3,
  specs = "political_party",
  by = "log_prev_week_inc_rate",
  at = list(log_prev_week_inc_rate = log_prev_week_inc_rate_values,
    wave = "2",
            interview_day_weekend = "Weekday",
            city = "National",
            age_group = "[18,30)",
            r_race = "White",
            resp_hispanic = "No",
            r_male = "0",
            r_working = "Working",
            resp_hhsize = average_hhsize,
            educ_group = "College graduate and above",
            county_mask_mandate = "Strict",
            r_metro = "1",
            CD_PERCENT_DEMOCRAT = average_CD_dem)
)


df_emm_contacts1 <- as.data.frame(summary(emm_contacts1))
df_emm_contacts2 <- as.data.frame(summary(emm_contacts2))
df_emm_contacts3 <- as.data.frame(summary(emm_contacts3))

# Add model identifiers
df_emm_contacts1$model <- "1"
df_emm_contacts2$model <- "2"
df_emm_contacts3$model <- "3"

# Filter out "Prefer not to answer"
df_emm_contacts1 <- df_emm_contacts1 %>% 
  filter(political_party != "Prefer not to answer")

df_emm_contacts2 <- df_emm_contacts2 %>% 
  filter(political_party != "Prefer not to answer")

df_emm_contacts3 <- df_emm_contacts3 %>% 
  filter(political_party != "Prefer not to answer")

all_emms <- rbind(df_emm_contacts1, df_emm_contacts2, df_emm_contacts3)

rvd_contacts_interaction_plot <- ggplot(all_emms, aes(x = log_prev_week_inc_rate, y = emmean, color = political_party, fill = political_party)) +
  geom_line() +
  geom_ribbon(aes(ymin = emmean - SE, ymax = emmean + SE), alpha = 0.5, linetype = "blank") +
  facet_wrap(~ model) +
  scale_fill_manual(values = c("Independent" = "#6db753", "Republican" = "#c50000", "Democrat" = "#326ba4")) +
  labs(
    title = "Impact on Daily Contacts",
    x = "Log Prev Week Inc Rate",
    y = "Estimated Marginal Means",
    color = "Political Party",
    fill = "Political Party"
  ) +
  theme_cowplot() +
  theme(
    strip.text = element_blank(),
    plot.title = element_text(size = 22),
    axis.title.x = element_blank(),
    axis.text.x = element_text(size = 15),
    axis.text.y = element_text(size = 22),
    axis.title.y = element_blank(),
    legend.title = element_text(size = 12),
    legend.text = element_text(size = 16),
    panel.grid.major.y = element_line(color = "grey", size = 0.5)
  )

rvd_contacts_interaction_plot

#version two plot, that shows ten percent increase in log prev week incidence rate
mean_log_prev_week_inc_rate <- mean(log_prev_week_inc_rate_values, na.rm = TRUE)
log_prev_week_inc_rate_values_10pct <- mean_log_prev_week_inc_rate * 1.10

# Calculate estimated means for mean values
emm_contacts1 <- emmeans(
  rvd_contacts_interaction_model1,
  specs = "political_party",
  by = "log_prev_week_inc_rate",
  at = list(log_prev_week_inc_rate = mean_log_prev_week_inc_rate,
            wave = "2",
            interview_day_weekend = "Weekday",
            city = "National")
)

# Calculate estimated means for increased values
emm_contacts1_10pct <- emmeans(
  rvd_contacts_interaction_model1,
  specs = "political_party",
  by = "log_prev_week_inc_rate",
  at = list(log_prev_week_inc_rate = log_prev_week_inc_rate_values_10pct,
            wave = "2",
            interview_day_weekend = "Weekday",
            city = "National"
))

combined_emm <- rbind(
  increased_value = emm_contacts1_10pct,
  mean_value = emm_contacts1
)

contrast_results_reversed <- contrast(
  combined_emm,
  method = "pairwise",
  by = "political_party",
  reverse = TRUE
)

emm_diff_reversed_inc <- as.data.frame(contrast_results_reversed)

emm_diff_reversed_inc <- emm_diff_reversed_inc %>% 
  filter(political_party != "Prefer not to answer")


rvd_contacts_interaction_diff_plot <- ggplot(emm_diff_reversed_inc, aes(x = political_party, y = estimate, fill = political_party)) +
  geom_bar(stat = "identity", alpha = 0.7) +
  scale_fill_manual(values = c("#326ba4", "#6db753", "#c50000")) +
  geom_errorbar(aes(ymin = estimate - SE, ymax = estimate + SE), width = 0.2, color = "black") +
  labs(
    title = "Contacts in Response to Ten Percent Increase in Previous Week Incidence",
    x = "Political Party",
    y = "Change in Contacts"
  ) +
  ylim(-1.1, 0.25) +
  theme_cowplot() +
  theme(
    strip.text = element_blank(),
    plot.title = element_text(size = 22),
    axis.title.x = element_blank(),
    axis.text.x = element_text(size = 15),
    axis.text.y = element_text(size = 22),
    axis.title.y = element_blank(),
    legend.position = "none", 
    panel.grid.major.y = element_line(color = "grey", size = 0.5)
  )

print(rvd_contacts_interaction_diff_plot)
```

# =============================================================================
# 6. CONTACTS RESPONSE TO MORTALITY (BY PARTY)
# =============================================================================
```{r}
rvd_contacts_interaction_model1 <- lm(num_cc ~ political_party*log_prev_week_mort_rate + wave + interview_day_weekend + city,
                          data = bics_zip_features,
                          weights = weight_party_raked)

print(summary(rvd_contacts_interaction_model1))

rvd_contacts_interaction_model2 <- lm(num_cc ~ political_party*log_prev_week_mort_rate + wave + interview_day_weekend + city + age_group + r_race + resp_hispanic + r_male + educ_group + r_working + resp_hhsize, 
                                   data = bics_zip_features, 
                                   weights = weight_party_raked)


print(summary(rvd_contacts_interaction_model2))


rvd_contacts_interaction_model3 <- lm(num_cc ~ political_party*log_prev_week_mort_rate + county_mask_mandate + r_metro + male + educ_group + age_group + resp_hhsize + r_race + resp_hispanic + r_working + city + wave + interview_day_weekend + CD_PERCENT_DEMOCRAT + STATE, 
                                   data = bics_zip_features, 
                                   weights = weight_party_raked)

print(summary(rvd_contacts_interaction_model3))


log_prev_week_mort_rate_values <- seq(from = min(bics_zip_features$log_prev_week_mort_rate, na.rm = TRUE),
                                     to = max(bics_zip_features$log_prev_week_mort_rate, na.rm = TRUE),
                                     length.out = 100)

emm_contacts1 <- emmeans(
  rvd_contacts_interaction_model1,
  specs = "political_party",
  by = "log_prev_week_mort_rate",
  at = list(log_prev_week_mort_rate = log_prev_week_mort_rate_values,
            wave = "2",
            interview_day_weekend = "Weekday",
            city = "National")
)


average_yob <- weighted.mean(bics_zip_features$age_group, w = bics_zip_features$weight_party_raked, na.rm = TRUE)
average_hhsize <- weighted.mean(bics_zip_features$resp_hhsize, w = bics_zip_features$weight_party_raked, na.rm = TRUE)

emm_contacts2 <- emmeans(
  rvd_contacts_interaction_model2,
  specs = "political_party",
  by = "log_prev_week_mort_rate",
  at = list(log_prev_week_mort_rate = log_prev_week_mort_rate_values,
            wave = "2",
            interview_day_weekend = "Weekday",
            city = "National",
            age_group = "[18,30)",
            r_race = "White",
            resp_hispanic = "Yes",
            r_male = "0",
            r_working = "Working",
            resp_hhsize = average_hhsize,
            educ_group = "College graduate and above")
)


average_CD_dem <- weighted.mean(bics_zip_features$CD_PERCENT_DEMOCRAT, w = bics_zip_features$weight_party_raked, na.rm = TRUE)

emm_contacts3 <- emmeans(
  rvd_contacts_interaction_model3,
  specs = "political_party",
  by = "log_prev_week_mort_rate",
  at = list(log_prev_week_mort_rate = log_prev_week_mort_rate_values,
    wave = "2",
            interview_day_weekend = "Weekday",
            city = "National",
            age_group = "[18,30)",
            r_race = "White",
            resp_hispanic = "Yes",
            r_male = "0",
            r_working = "Working",
            resp_hhsize = average_hhsize,
            educ_group = "College graduate and above",
            county_mask_mandate = "Strict",
            r_metro = "1",
            CD_PERCENT_DEMOCRAT = average_CD_dem)
)


df_emm_contacts1 <- as.data.frame(summary(emm_contacts1))
df_emm_contacts2 <- as.data.frame(summary(emm_contacts2))
df_emm_contacts3 <- as.data.frame(summary(emm_contacts3))

# Add model identifiers
df_emm_contacts1$model <- "1"
df_emm_contacts2$model <- "2"
df_emm_contacts3$model <- "3"

# Filter out "Prefer not to answer"
df_emm_contacts1 <- df_emm_contacts1 %>% 
  filter(political_party != "Prefer not to answer")

df_emm_contacts2 <- df_emm_contacts2 %>% 
  filter(political_party != "Prefer not to answer")

df_emm_contacts3 <- df_emm_contacts3 %>% 
  filter(political_party != "Prefer not to answer")

all_emms <- rbind(df_emm_contacts1, df_emm_contacts2, df_emm_contacts3)

rvd_contacts_interaction_plot_mort <- ggplot(all_emms, aes(x = log_prev_week_mort_rate, y = emmean, color = political_party, fill = political_party)) +
  geom_line() +
  geom_ribbon(aes(ymin = emmean - SE, ymax = emmean + SE), alpha = 0.5, linetype = "blank") +
  facet_wrap(~ model) +
  scale_fill_manual(values = c("Independent" = "#6db753", "Republican" = "#c50000", "Democrat" = "#326ba4")) +
  labs(
    title = "Impact on Daily Contacts",
    x = "Log Mort Week Inc Rate",
    y = "Estimated Marginal Means",
    color = "Political Party",
    fill = "Political Party"
  ) +
  theme_cowplot() +
  theme(
    strip.text = element_blank(),
    plot.title = element_text(size = 22),
    axis.title.x = element_blank(),
    axis.text.x = element_text(size = 15),
    axis.text.y = element_text(size = 22),
    axis.title.y = element_blank(),
    legend.title = element_text(size = 12),
    legend.text = element_text(size = 16),
    panel.grid.major.y = element_line(color = "grey", size = 0.5)
  )

rvd_contacts_interaction_plot_mort

#version two plot, that shows ten percent increase in log mort week incidence rate
mean_log_prev_week_mort_rate <- mean(log_prev_week_mort_rate_values, na.rm = TRUE)
log_prev_week_mort_rate_values_10pct <- mean_log_prev_week_mort_rate * 1.10

# Calculate estimated means for mean values
emm_contacts1 <- emmeans(
  rvd_contacts_interaction_model1,
  specs = "political_party",
  by = "log_prev_week_mort_rate",
  at = list(log_log_prev_week_mort_rate = mean_log_prev_week_mort_rate,
            wave = "2",
            interview_day_weekend = "Weekday",
            city = "National")
)

# Calculate estimated means for increased values
emm_contacts1_10pct <- emmeans(
  rvd_contacts_interaction_model1,
  specs = "political_party",
  by = "log_prev_week_mort_rate",
  at = list(log_prev_week_mort_rate = log_prev_week_mort_rate_values_10pct,
            wave = "2",
            interview_day_weekend = "Weekday",
            city = "National")
)

combined_emm <- rbind(
  increased_value = emm_contacts1_10pct,
  mean_value = emm_contacts1
)

contrast_results_reversed <- contrast(
  combined_emm,
  method = "pairwise",
  by = "political_party",
  reverse = TRUE
)

emm_diff_reversed_mort <- as.data.frame(contrast_results_reversed)

emm_diff_reversed_mort <- emm_diff_reversed_mort %>% 
  filter(political_party != "Prefer not to answer")


rvd_contacts_interaction_diff_plot_mort <- ggplot(emm_diff_reversed_mort, aes(x = political_party, y = estimate, fill = political_party)) +
  geom_bar(stat = "identity", alpha = 0.7, width = 0.7) +  # Reduced width for better visibility of error bars
  scale_fill_manual(values = c("#326ba4", "#6db753", "#c50000")) +
  geom_errorbar(aes(ymin = estimate - SE, ymax = estimate + SE), width = 0.2, color = "black", position = position_dodge(width = 0.7)) +
  labs(
    title = "Contacts in Response to Ten Percent Increase in Previous Week Mortality",
    x = "Political Party",
    y = "Change in Contacts"
  ) +
  coord_cartesian(ylim = c(-1.1, 0.25)) +  # Use coord_cartesian instead of ylim
  theme_cowplot() +
  theme(
    strip.text = element_blank(),
    plot.title = element_text(size = 22),
    axis.title.x = element_blank(),
    axis.text.x = element_text(size = 15),
    axis.text.y = element_text(size = 22),
    axis.title.y = element_blank(),
    legend.position = "none", 
    panel.grid.major.y = element_line(color = "grey", size = 0.5)
  )

print(rvd_contacts_interaction_diff_plot_mort)
```

# =============================================================================
# 7. COMBINED: CONTACTS BY PARTY
# =============================================================================
```{r}
rvd_contacts_interaction_diff_plot_mort <- rvd_contacts_interaction_diff_plot_mort + 
  theme(legend.position = "none")

# Combine plots without legends
combined_plot <- plot_grid(
  plot_grid(rvd_contacts_interaction_diff_plot + theme(legend.position = "none"),
            rvd_contacts_interaction_diff_plot_mort,
            ncol = 1, labels = c("A", "B")),
  ncol = 1
)

# Add a label to the combined plot
combined_plot <- plot_grid(
  ggdraw() + draw_label("Figure", size = 16, hjust = 0, x = 0.01),
  combined_plot,
  ncol = 1,
  rel_heights = c(0.4, 10)
)

# Display the final combined plot
print(combined_plot)

ggsave("combined_national_pooled_mortalty_and_incidence_response_contacts.png", plot = combined_plot, path = "../out/plots", width = 14, height = 16, units = "in")
```

# =============================================================================
# 8. CONTACTS BY WAVE AND MORTALITY
# =============================================================================

I want to create a plot that shows predicted contacts based on average mortality rate at each wave.
```{r}
wave_avgs <- list()

for (i in 1:6) {
  avg <- bics_zip_features %>%
    filter(wave == i) %>%
    summarise(avg_log_prev_week_mort_rate = mean(log_prev_week_mort_rate, na.rm = TRUE)) %>%
    pull(avg_log_prev_week_mort_rate)
  
  wave_avgs[[paste0("wave", i)]] <- avg
}

print(wave_avgs)

wave1_estimated_contacts <- emmeans(
  rvd_contacts_interaction_model1,
  specs = "political_party",
  by = "log_prev_week_mort_rate",
  at = list(log_prev_week_mort_rate = wave_avgs$wave1,
            wave = "1",
            interview_day_weekend = "Weekday",
            city = "National")
)

wave_estimated_contacts <- list()

for (i in 1:6) {
  wave_name <- paste0("wave", i)
  
  estimated_contacts <- emmeans(
    rvd_contacts_interaction_model1,
    specs = "political_party",
    by = "log_prev_week_mort_rate",
    at = list(log_prev_week_mort_rate = wave_avgs[[wave_name]],
              wave = "2",
              interview_day_weekend = "Weekday",
              city = "National")
  )
  
  wave_estimated_contacts[[wave_name]] <- estimated_contacts
  
}

combined_estimated_contacts <- data.frame()

for (i in 1:6) {
  wave_name <- paste0("wave", i)
  
  wave_df <- as.data.frame(wave_estimated_contacts[[wave_name]])
  
  wave_df$wave <- i
  
  combined_estimated_contacts <- rbind(combined_estimated_contacts, wave_df)
}

rownames(combined_estimated_contacts) <- NULL

print(head(combined_estimated_contacts))

combined_estimated_contacts <- combined_estimated_contacts %>%
  filter(political_party != "Prefer not to answer")

contacts_mortality_response_waves <- ggplot(combined_estimated_contacts, aes(x = wave, y = emmean, color = political_party, group = political_party)) +
  geom_point(size = 3, position = position_dodge(width = 0.3), alpha = 0.7) +
  geom_errorbar(aes(ymin = emmean - SE, ymax = emmean + SE), width = 0.2, position = position_dodge(width = 0.3), color = "black", alpha = 0.7) +
  geom_line(position = position_dodge(width = 0.3)) +
  scale_x_continuous(breaks = 1:6) +
  scale_color_manual(values = c("Democrat" = "#326ba4", "Independent" = "#6db753", "Republican" = "#c50000")) +
  labs(title = "Estimated Marginal Means of Contacts by Political Party and County-Level \nMortality Rates",
       x = "Average Mortality Rate at Time of Wave",
       y = "Estimated Mean Contacts",
       color = "Political Party") +
  theme(legend.position = "bottom",
        panel.grid.minor = element_blank()) +
  theme_cowplot() +
  theme(
    strip.text = element_blank(),
    plot.title = element_text(size = 22),
    axis.title.x = element_text(size = 13),
    axis.text.x = element_text(size = 15),
    axis.text.y = element_text(size = 22),
    axis.title.y = element_blank(),
    legend.position = "bottom", 
    panel.grid.major.y = element_line(color = "grey", size = 0.5)
  )

contacts_mortality_response_waves

ggsave("contacts_mortality_response_waves.png", plot = contacts_mortality_response_waves, path = "../out/plots", width = 12, height = 8, units = "in")
```

# =============================================================================
# 9. MASKS RESPONSE TO INCIDENCE (BY PARTY)
# =============================================================================
```{r}
rvd_masks_interaction_model1 <- lm(Norm_Masks_Used ~ political_party*log_prev_week_inc_rate + wave + interview_day_weekend + city,
                                   data = bics_zip_features,
                                   weights = weight_party_raked)

rvd_masks_interaction_model2 <- lm(Norm_Masks_Used ~ political_party*log_prev_week_inc_rate + wave + interview_day_weekend + city + age_group + r_race + resp_hispanic + r_male + educ_group + r_working + resp_hhsize, 
                                   data = bics_zip_features, 
                                   weights = weight_party_raked)

rvd_masks_interaction_model3 <- lm(Norm_Masks_Used ~ political_party*log_prev_week_inc_rate + county_mask_mandate + r_metro + male + educ_group + age_group + resp_hhsize + r_race + resp_hispanic + r_working + city + wave + interview_day_weekend + CD_PERCENT_DEMOCRAT + STATE, 
                                   data = bics_zip_features, 
                                   weights = weight_party_raked)

mean_log_prev_week_inc_rate <- mean(log_prev_week_inc_rate_values, na.rm = TRUE)
log_prev_week_inc_rate_values_10pct <- mean_log_prev_week_inc_rate * 1.10

emm_masks1 <- emmeans(
  rvd_masks_interaction_model1,
  specs = "political_party",
  by = "log_prev_week_inc_rate",
  at = list(log_prev_week_inc_rate = log_prev_week_inc_rate_values,
            wave = "2",
            interview_day_weekend = "Weekday",
            city = "National")
)

emm_masks2 <- emmeans(
  rvd_masks_interaction_model2,
  specs = "political_party",
  by = "log_prev_week_inc_rate",
  at = list(log_prev_week_inc_rate = log_prev_week_inc_rate_values,
            wave = "2",
            interview_day_weekend = "Weekday",
            city = "National",
            age_group = "[18,30)",
            r_race = "White",
            resp_hispanic = "Yes",
            r_male = "0",
            r_working = "Working",
            resp_hhsize = average_hhsize,
            educ_group = "College graduate and above")
)

emm_masks3 <- emmeans(
  rvd_masks_interaction_model3,
  specs = "political_party",
  by = "log_prev_week_inc_rate",
  at = list(log_prev_week_inc_rate = log_prev_week_inc_rate_values,
            wave = "2",
            interview_day_weekend = "Weekday",
            city = "National",
            age_group = "[18,30)",
            r_race = "White",
            resp_hispanic = "Yes",
            r_male = "0",
            r_working = "Working",
            resp_hhsize = average_hhsize,
            educ_group = "College graduate and above",
            county_mask_mandate = "Strict",
            r_metro = "1",
            CD_PERCENT_DEMOCRAT = average_CD_dem)
)

df_emm_masks1 <- as.data.frame(summary(emm_masks1))
df_emm_masks2 <- as.data.frame(summary(emm_masks2))
df_emm_masks3 <- as.data.frame(summary(emm_masks3))

df_emm_masks1$model <- "1"
df_emm_masks2$model <- "2"
df_emm_masks3$model <- "3"

# Filter out "Prefer not to answer"
df_emm_masks1 <- df_emm_masks1 %>% 
  filter(political_party != "Prefer not to answer")

df_emm_masks2 <- df_emm_masks2 %>% 
  filter(political_party != "Prefer not to answer")

df_emm_masks3 <- df_emm_masks3 %>% 
  filter(political_party != "Prefer not to answer")

# Combine data frames
all_emms_masks <- rbind(df_emm_masks1, df_emm_masks2, df_emm_masks3)

# Plot
rvd_masks_interaction_plot <- ggplot(all_emms_masks, aes(x = log_prev_week_inc_rate, y = emmean, color = political_party, fill = political_party)) +
  geom_line() +
  geom_ribbon(aes(ymin = emmean - SE, ymax = emmean + SE), alpha = 0.5, linetype = "blank") +
  facet_wrap(~ model) +
  scale_fill_manual(values = c("Independent" = "#6db753", "Republican" = "#c50000", "Democrat" = "#326ba4")) +
  labs(
    title = "Impact on Percent of Contacts With a Mask",
    x = "Log Prev Week Inc Rate",
    y = "Estimated Marginal Means",
    color = "Political Party",
    fill = "Political Party"
  ) +
  theme_cowplot() +
  theme(
    strip.text = element_blank(),
    plot.title = element_text(size = 22),
    axis.title.x = element_blank(),
    axis.text.x = element_text(size = 15),
    axis.text.y = element_text(size = 22),
    axis.title.y = element_blank(),
    legend.title = element_text(size = 12),
    legend.text = element_text(size = 16),
    panel.grid.major.y = element_line(color = "grey", size = 0.5)
  )

rvd_masks_interaction_plot

mean_log_prev_week_mort_rate <- mean(log_prev_week_mort_rate_values, na.rm = TRUE)
log_prev_week_mort_rate_values_10pct <- mean_log_prev_week_mort_rate * 1.10

emm_masks1 <- emmeans(
  rvd_masks_interaction_model1,
  specs = "political_party",
  by = "log_prev_week_inc_rate",
  at = list(log_prev_week_inc_rate = mean_log_prev_week_inc_rate,
            wave = "2",
            interview_day_weekend = "Weekday",
            city = "National")
)

# Calculate estimated means for increased values
emm_masks1_10pct <- emmeans(
  rvd_masks_interaction_model1,
  specs = "political_party",
  by = "log_prev_week_inc_rate",
  at = list(log_prev_week_inc_rate = log_prev_week_inc_rate_values_10pct,
            wave = "2",
            interview_day_weekend = "Weekday",
            city = "National"
))

combined_emm <- rbind(
  increased_value = emm_masks1_10pct,
  mean_value = emm_masks1
)

contrast_results_reversed <- contrast(
  combined_emm,
  method = "pairwise",
  by = "political_party",
  reverse = TRUE
)

emm_diff_reversed_inc_masks <- as.data.frame(contrast_results_reversed)

emm_diff_reversed_inc_masks <- emm_diff_reversed_inc_masks %>% 
  filter(political_party != "Prefer not to answer")

rvd_masks_interaction_diff_plot_inc <- ggplot(emm_diff_reversed_inc_masks, aes(x = political_party, y = estimate, fill = political_party)) +
  geom_bar(stat = "identity", alpha = 0.7) +
  geom_errorbar(aes(ymin = estimate - SE, ymax = estimate + SE), width = 0.2, color = "black") +
  scale_fill_manual(values = c("#326ba4", "#6db753", "#c50000")) +
  labs(
    title = "Mask Usage in Response to 10% Increase in Previous Week Incidence",
    x = "Political Party",
    y = "Change in Mask Usage"
  ) +
  theme_cowplot() +
  theme(
    strip.text = element_blank(),
    plot.title = element_text(size = 22),
    axis.title.x = element_blank(),
    axis.text.x = element_text(size = 15),
    axis.text.y = element_text(size = 22),
    axis.title.y = element_blank(),
    legend.position = "none", 
    panel.grid.major.y = element_line(color = "grey", size=0.5)
  )

print(rvd_masks_interaction_diff_plot_inc)
```

# =============================================================================
# 10. MASKS RESPONSE TO MORTALITY (BY PARTY)
# =============================================================================
```{r}
rvd_masks_interaction_model1 <- lm(Norm_Masks_Used ~ political_party*log_prev_week_mort_rate + wave + interview_day_weekend + city,
                                   data = bics_zip_features,
                                   weights = weight_party_raked)

rvd_masks_interaction_model2 <- lm(Norm_Masks_Used ~ political_party*log_prev_week_mort_rate + wave + interview_day_weekend + city + age_group + r_race + resp_hispanic + r_male + educ_group + r_working + resp_hhsize, 
                                   data = bics_zip_features, 
                                   weights = weight_party_raked)

rvd_masks_interaction_model3 <- lm(Norm_Masks_Used ~ political_party*log_prev_week_mort_rate + county_mask_mandate + r_metro + male + educ_group + age_group + resp_hhsize + r_race + resp_hispanic + r_working + city + wave + interview_day_weekend + CD_PERCENT_DEMOCRAT + STATE, 
                                   data = bics_zip_features, 
                                   weights = weight_party_raked)

mean_log_prev_week_mort_rate <- mean(log_prev_week_mort_rate_values, na.rm = TRUE)
log_prev_week_mort_rate_values_10pct <- mean_log_prev_week_mort_rate * 1.10

emm_masks1 <- emmeans(
  rvd_masks_interaction_model1,
  specs = "political_party",
  by = "log_prev_week_mort_rate",
  at = list(log_prev_week_mort_rate = log_prev_week_mort_rate_values,
            wave = "2",
            interview_day_weekend = "Weekday",
            city = "National")
)

emm_masks2 <- emmeans(
  rvd_masks_interaction_model2,
  specs = "political_party",
  by = "log_prev_week_mort_rate",
  at = list(log_prev_week_mort_rate = log_prev_week_mort_rate_values,
            wave = "2",
            interview_day_weekend = "Weekday",
            city = "National",
            age_group = "[18,30)",
            r_race = "White",
            resp_hispanic = "Yes",
            r_male = "0",
            r_working = "Working",
            resp_hhsize = average_hhsize,
            educ_group = "College graduate and above")
)

emm_masks3 <- emmeans(
  rvd_masks_interaction_model3,
  specs = "political_party",
  by = "log_prev_week_mort_rate",
  at = list(log_prev_week_mort_rate = log_prev_week_mort_rate_values,
            wave = "2",
            interview_day_weekend = "Weekday",
            city = "National",
            age_group = "[18,30)",
            r_race = "White",
            resp_hispanic = "Yes",
            r_male = "0",
            r_working = "Working",
            resp_hhsize = average_hhsize,
            educ_group = "College graduate and above",
            county_mask_mandate = "Strict",
            r_metro = "1",
            CD_PERCENT_DEMOCRAT = average_CD_dem)
)

df_emm_masks1 <- as.data.frame(summary(emm_masks1))
df_emm_masks2 <- as.data.frame(summary(emm_masks2))
df_emm_masks3 <- as.data.frame(summary(emm_masks3))

df_emm_masks1$model <- "1"
df_emm_masks2$model <- "2"
df_emm_masks3$model <- "3"

# Filter out "Prefer not to answer"
df_emm_masks1 <- df_emm_masks1 %>% 
  filter(political_party != "Prefer not to answer")

df_emm_masks2 <- df_emm_masks2 %>% 
  filter(political_party != "Prefer not to answer")

df_emm_masks3 <- df_emm_masks3 %>% 
  filter(political_party != "Prefer not to answer")

# Combine data frames
all_emms_masks <- rbind(df_emm_masks1, df_emm_masks2, df_emm_masks3)

# Plot
rvd_masks_interaction_plot_mort <- ggplot(all_emms_masks, aes(x = log_prev_week_mort_rate, y = emmean, color = political_party, fill = political_party)) +
  geom_line() +
  geom_ribbon(aes(ymin = emmean - SE, ymax = emmean + SE), alpha = 0.5, linetype = "blank") +
  facet_wrap(~ model) +
  scale_fill_manual(values = c("Independent" = "#6db753", "Republican" = "#c50000", "Democrat" = "#326ba4")) +
  labs(
    title = "Impact on Percent of Contacts With a Mask",
    x = "Log Prev Week Mort Rate",
    y = "Estimated Marginal Means",
    color = "Political Party",
    fill = "Political Party"
  ) +
  theme_cowplot() +
  theme(
    strip.text = element_blank(),
    plot.title = element_text(size = 22),
    axis.title.x = element_blank(),
    axis.text.x = element_text(size = 15),
    axis.text.y = element_text(size = 22),
    axis.title.y = element_blank(),
    legend.title = element_text(size = 12),
    legend.text = element_text(size = 16),
    panel.grid.major.y = element_line(color = "grey", size = 0.5)
  )

rvd_masks_interaction_plot_mort

rvd_masks_interaction_plot

mean_log_prev_week_mort_rate <- mean(log_prev_week_mort_rate_values, na.rm = TRUE)
log_prev_week_mort_rate_values_10pct <- mean_log_prev_week_mort_rate * 1.10

emm_masks1 <- emmeans(
  rvd_masks_interaction_model1,
  specs = "political_party",
  by = "log_prev_week_mort_rate",
  at = list(log_prev_week_mort_rate = mean_log_prev_week_mort_rate,
            wave = "2",
            interview_day_weekend = "Weekday",
            city = "National")
)

# Calculate estimated means for increased values
emm_masks1_10pct <- emmeans(
  rvd_masks_interaction_model1,
  specs = "political_party",
  by = "log_prev_week_mort_rate",
  at = list(log_prev_week_mort_rate = log_prev_week_mort_rate_values_10pct,
            wave = "2",
            interview_day_weekend = "Weekday",
            city = "National"
))

combined_emm <- rbind(
  increased_value = emm_masks1_10pct,
  mean_value = emm_masks1
)

contrast_results_reversed <- contrast(
  combined_emm,
  method = "pairwise",
  by = "political_party",
  reverse = TRUE
)

emm_diff_reversed_mort_masks <- as.data.frame(contrast_results_reversed)

emm_diff_reversed_mort_masks <- emm_diff_reversed_mort_masks %>% 
  filter(political_party != "Prefer not to answer")

rvd_masks_interaction_diff_plot_mort <- ggplot(emm_diff_reversed_mort_masks, aes(x = political_party, y = estimate, fill = political_party)) +
  geom_bar(stat = "identity", alpha = 0.7) +
  geom_errorbar(aes(ymin = estimate - SE, ymax = estimate + SE), width = 0.2, color = "black") +
  scale_fill_manual(values = c("#326ba4", "#6db753", "#c50000")) +
  labs(
    title = "Mask Usage in Response to 10% Increase in Previous Week Mortality",
    x = "Political Party",
    y = "Change in Mask Usage"
  ) +
  theme_cowplot() +
  theme(
    strip.text = element_blank(),
    plot.title = element_text(size = 22),
    axis.title.x = element_blank(),
    axis.text.x = element_text(size = 15),
    axis.text.y = element_text(size = 22),
    axis.title.y = element_blank(),
    legend.position = "none", 
    panel.grid.major.y = element_line(color = "grey", size=0.5)
  )

print(rvd_masks_interaction_diff_plot_mort)
```

# =============================================================================
# 11. COMBINED: MASKS BY PARTY
# =============================================================================
```{r}
rvd_masks_interaction_diff_plot_mort <- rvd_masks_interaction_diff_plot_mort + 
  theme(legend.position = "none")

# Combine plots without legends
combined_plot <- plot_grid(
  plot_grid(rvd_masks_interaction_diff_plot_inc + theme(legend.position = "none"),
            rvd_masks_interaction_diff_plot_mort,
            ncol = 1, labels = c("A", "B")),
  ncol = 1
)

# Add a label to the combined plot
combined_plot <- plot_grid(
  combined_plot,
  ncol = 1,
  rel_heights = c(0.4, 10)
)

# Display the final combined plot
print(combined_plot)

ggsave("combined_national_pooled_mortalty_and_incidence_response_masks.png", plot = combined_plot, path = "../out/plots", width = 14, height = 16, units = "in")
```

# =============================================================================
# 12. CLEANUP
# =============================================================================

```{r cleanup}
rm(rvd_vaccinated_interaction_plot, rvd_vaccinated_interaction_model3, rvd_vaccinated_interaction_model2, rvd_vaccinated_interaction_model1, rvd_masks_interaction_diff_plot, rvd_masks_interaction_plot, all_emms_vaccinated, all_emms, all_emms_masks, combined_emm, combined_plot, contrast_results_reversed, emm_vaccinated3, emm_vaccinated2, emm_vaccinated1, emm_vaccination3, emm_vaccination2, emm_vaccination1, emm_diff_reversed, log_prev_week_inc_rate_values, log_prev_week_inc_rate_values_10pct, log_prev_week_mort_rate_values, log_prev_week_mort_rate_values_10pct, average_CD_dem, average_hhsize, average_yob, mean_log_prev_week_inc_rate, mean_log_prev_week_mort_rate, mean_log_prev_week_mort_rate_rate, df_emm_contacts1, df_emm_contacts2, df_emm_contacts3, df_emm_masks1, df_emm_masks2, df_emm_masks3, emm_contacts1, emm_contacts1_10pct, emm_contacts2, emm_contacts3, emm_diff_reversed_inc, emm_diff_reversed_inc_masks, emm_diff_reversed_mort, emm_diff_reversed_mort_masks, emm_masks1, emm_masks1_10pct, emm_masks2, emm_masks3, rvd_contacts_interaction_diff_plot, rvd_contacts_interaction_diff_plot_mort, rvd_contacts_interaction_model1, rvd_contacts_interaction_model2, rvd_contacts_interaction_model3, rvd_masks_interaction_plot, rvd_contacts_interaction_plot, rvd_contacts_interaction_plot_mort, rvd_masks_interaction_diff_plot_inc, rvd_masks_interaction_diff_plot_mort, rvd_masks_interaction_model1, rvd_masks_interaction_model2, rvd_masks_interaction_model3, rvd_masks_interaction_plot_mort, combined_estimated_contacts, contacts_interaction_diff_plot, contacts_interaction_diff_plot_mort, contacts_interaction_model1, contacts_interaction_model2, contacts_interaction_model3, contacts_interaction_plot, contacts_mortality_response_waves, contrast_df, contrast_results, wave_avgs, R_county_CD_share, wave_avgs, wave_estimated_contacts, wave1_estimated_contacts, avg, dodge_width, i, wave_name, wave_df, estimated_contacts)
```

