---
editor: 
  markdown: 
    wrap: sentence
---

# Results {#sec-results}

```{r}
#| label: results-setup
#| include: false

# Load pre-calculated statistics from the export script
# Run code/export_paper_statistics.R to regenerate this file
stats_path <- "../out/tables/paper_statistics.rds"

if (!file.exists(stats_path)) {
  stop("Paper statistics file not found. Please run code/export_paper_statistics.R first.")
}

# Load statistics and unpack into environment
paper_stats <- readRDS(stats_path)
list2env(paper_stats, envir = environment())
```

We structure our results as a series of tests designed to adjudicate between the competing theoretical explanations outlined in the introduction.
Each section confronts a specific theory with the empirical patterns: if the theory is correct, what should we observe—and do we?
We first establish that partisan differences in health behaviors exist and are substantial, surviving demographic controls that would eliminate them if they were artifacts of compositional differences.
We then test whether partisan gaps in risk perception can account for behavioral differences, and whether objective pandemic severity overrides partisan identity when the stakes are highest.
Finally, we assess whether geographic context, policy environments, or local partisan composition can explain the gaps—or whether partisan differences persist even in the conditions where each theory predicts they should disappear.

## Establishing Partisan Differences in Health Behaviors {#sec-behaviors}

We find substantial partisan differences across all three health behaviors examined: social contacts, mask-wearing, and vaccination.
These differences were consistent across survey waves and persisted even after controlling for demographic characteristics including age, gender, race, education, and urban/rural residence — eliminating the possibility that partisan gaps are artifacts of demographic sorting (the "it's just demographics" explanation).
Republicans reported the highest levels of social contact, the lowest rates of mask usage, and the lowest vaccine acceptance rates, while Democrats exhibited the opposite pattern.
Independents generally fell between the two partisan groups, though often closer to Republicans on contact behavior and closer to Democrats on vaccine acceptance.
@fig-partisan-behaviors provides a visual summary of these differences.

```{r}
#| label: fig-partisan-behaviors
#| fig-cap: "Partisan Differences in Health Behaviors. Panel A shows mean daily contacts, Panel B shows percentage of contacts during which masks were worn, and Panel C shows vaccine acceptance rates (Wave 6 only, May 2021). Vaccine acceptance includes those who were vaccinated or intended to get vaccinated. Values are weighted means using raked survey weights."
#| fig-height: 4
#| fig-width: 10

library(cowplot)

# Party colors (matching other BICS plots)
party_colors <- c("Democrat" = "#326ba4", "Independent" = "#6db753", "Republican" = "#c50000")

# Create data for plotting
behavior_data <- tibble(
  party = rep(c("Democrat", "Independent", "Republican"), 3),
  behavior = c(rep("A. Daily Contacts", 3),
               rep("B. Mask Usage (%)", 3),
               rep("C. Vaccine Acceptance (%)", 3)),
  value = c(contacts_dem, contacts_ind, contacts_rep,
            mask_dem, mask_ind, mask_rep,
            vax_dem, vax_ind, vax_rep)
) %>%
  mutate(
    party = factor(party, levels = c("Republican", "Independent", "Democrat")),
    behavior = factor(behavior, levels = c("A. Daily Contacts", "B. Mask Usage (%)", "C. Vaccine Acceptance (%)"))
  )

# Create faceted bar plot with cowplot styling
ggplot(behavior_data, aes(x = party, y = value, fill = party)) +
  geom_col(alpha = 0.7, width = 0.7, color = "black") +
  geom_text(aes(label = sprintf("%.1f", value)),
            vjust = -0.5, size = 3.2, fontface = "bold") +
  facet_wrap(~ behavior, scales = "free_y", ncol = 3) +
  scale_fill_manual(values = party_colors) +
  labs(x = NULL, y = NULL) +
  theme_cowplot() +
  theme(
    legend.position = "none",
    axis.text.x = element_text(angle = 0, hjust = 0.5, size = 10),
    strip.text = element_text(face = "bold", size = 11),
    strip.background = element_blank()
  ) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.15)))
```

### Social Contacts

Republicans reported significantly more daily interpersonal contacts than Democrats throughout the study period.
On average, Republicans reported `r sprintf("%.1f", contacts_rep)` total daily contacts compared to `r sprintf("%.1f", contacts_dem)` for Democrats, a difference of approximately `r sprintf("%.1f", contacts_gap)` contacts per day.
This gap was even more pronounced for non-household contacts, with Republicans averaging `r sprintf("%.1f", nonhh_rep)` non-household contacts compared to `r sprintf("%.1f", nonhh_dem)` for Democrats.
Independents fell between the two groups, reporting an average of `r sprintf("%.1f", contacts_ind)` total daily contacts and `r sprintf("%.1f", nonhh_ind)` non-household contacts.
These differences in contact patterns have important implications for disease transmission, as higher contact rates increase opportunities for viral spread within social networks.

### Mask-Wearing

Partisan differences in mask-wearing were substantial and represented one of the largest behavioral gaps observed in the study.
Democrats reported wearing masks during `r sprintf("%.1f", mask_dem)`% of their interpersonal contacts, compared to just `r sprintf("%.1f", mask_rep)`% for Republicans---a gap of nearly `r sprintf("%.0f", mask_gap)` percentage points.
Independents reported mask usage rates (`r sprintf("%.1f", mask_ind)`%) similar to Republicans.
This partisan gap in mask-wearing was larger than differences observed across other demographic categories including race, gender, and education, highlighting the outsized role of political affiliation in shaping protective behavior during the pandemic.

Partisan differences in vaccine acceptance followed the same pattern (see @sec-vaccine-appendix); because vaccination was assessed only in Wave 6 and cannot be examined longitudinally or geographically with the same depth, the analyses that follow focus on contacts and mask-wearing.
Underlying all three behavioral differences were substantial partisan gaps in pandemic risk perception (@fig-perception in @sec-perception-appendix): `r sprintf("%.0f", concern_dem)`% of Democrats reported being "very concerned" about COVID-19 compared to `r sprintf("%.0f", concern_rep)`% of Republicans---a `r sprintf("%.0f", concern_gap)` percentage point gap documented extensively in prior work [@gadarian_partisanship_2021; @milosh_unmasking_2021; @kaushal_american_2022].
The critical question is whether these perception differences can fully account for the behavioral gaps, or whether partisan identity shapes behavior through channels beyond information processing.
We turn to that question next.

## Does Perceived Pandemic Severity Mitigate Partisan Behavioral Polarization? {#sec-perceptions}

Democrats were substantially more likely than Republicans to report being "very concerned" about COVID-19 throughout the study period (@fig-perception in @sec-perception-appendix).
Overall, `r sprintf("%.0f", concern_dem)`% of Democrats expressed high concern, compared to just `r sprintf("%.0f", concern_rep)`% of Republicans; Independents were closer to Republicans at `r sprintf("%.0f", concern_ind)`%.

### Stated Concern and Behavioral Gaps

If partisan differences in behavior are driven by differences in perceived pandemic severity, then partisans who express the same level of concern should behave similarly.
@fig-concern-gaps tests this by comparing the Republican-Democrat gap in contacts and mask usage within each level of stated concern (see @fig-concern-behavior-appendix for levels by party).

```{r}
#| label: fig-concern-gaps
#| fig-cap: "Partisan Gaps in Health Behaviors Within Each Level of Stated Concern. Panel A shows the absolute partisan gap in daily contacts; Panel B shows the absolute partisan gap in mask usage. For contacts, the partisan gap persists at all concern levels. For masks, the gap narrows substantially among the 'very concerned.' Sample sizes: Not at all concerned (n=1,127), Not very concerned (n=1,897), Somewhat concerned (n=5,894), Very concerned (n=9,168)."
#| fig-height: 4
#| fig-width: 10

library(patchwork)
library(cowplot)

# Load behaviors by concern level
behaviors_by_concern <- read_csv("../out/tables/behaviors_by_concern_level.csv", show_col_types = FALSE) %>%
  mutate(
    political_party = factor(political_party, levels = c("Republican", "Independent", "Democrat")),
    concern_level = factor(concern_level,
                           levels = c("Not at all concerned", "Not very concerned",
                                      "Somewhat concerned", "Very concerned"))
  )

# Calculate absolute gaps at each concern level
gaps_by_concern <- behaviors_by_concern %>%
  select(concern_level, political_party, contacts, mask_usage) %>%
  pivot_wider(names_from = political_party, values_from = c(contacts, mask_usage)) %>%
  mutate(
    contacts_gap = abs(contacts_Republican - contacts_Democrat),
    mask_gap = abs(mask_usage_Democrat - mask_usage_Republican)
  )

# Panel A: Contact gap by concern level
p_contact_gap <- ggplot(gaps_by_concern, aes(x = concern_level, y = contacts_gap)) +
  geom_col(fill = "#7570b3", alpha = 0.8, color = "black", width = 0.6) +
  geom_text(aes(label = sprintf("%.1f", contacts_gap)), vjust = -0.5, size = 4, fontface = "bold") +
  scale_y_continuous(expand = expansion(mult = c(0, 0.15))) +
  scale_x_discrete(labels = c("Not at all\nconcerned", "Not very\nconcerned",
                              "Somewhat\nconcerned", "Very\nconcerned")) +
  labs(x = NULL, y = "Partisan Gap (contacts/day)", title = "A. Contact Gap") +
  theme_cowplot() +
  theme(
    plot.title = element_text(face = "bold", size = 11),
    axis.text.x = element_text(size = 9)
  )

# Panel B: Mask gap by concern level
p_mask_gap <- ggplot(gaps_by_concern, aes(x = concern_level, y = mask_gap)) +
  geom_col(fill = "#1b9e77", alpha = 0.8, color = "black", width = 0.6) +
  geom_text(aes(label = sprintf("%.0f", mask_gap)), vjust = -0.5, size = 4, fontface = "bold") +
  scale_y_continuous(expand = expansion(mult = c(0, 0.15))) +
  scale_x_discrete(labels = c("Not at all\nconcerned", "Not very\nconcerned",
                              "Somewhat\nconcerned", "Very\nconcerned")) +
  labs(x = NULL, y = "Partisan Gap (pp)", title = "B. Mask Gap") +
  theme_cowplot() +
  theme(
    plot.title = element_text(face = "bold", size = 11),
    axis.text.x = element_text(size = 9)
  )

# Combine
p_contact_gap + p_mask_gap
```

For social contacts, the partisan gap persists at every level of concern: Republicans report more contacts than Democrats whether they are "not at all concerned" or "very concerned" about the pandemic.
Among the very concerned, Republicans still reported `r sprintf("%.1f", contacts_vc_rep)` daily contacts compared to `r sprintf("%.1f", contacts_vc_dem)` for Democrats, a gap of `r sprintf("%.1f", contacts_vc_gap)` contacts.
For mask usage, however, perceived severity does appear to mitigate polarization: the partisan gap narrows as concern increases, and among the "very concerned," Republicans and Democrats report nearly identical mask usage (`r sprintf("%.0f", mask_vc_rep)`% vs. `r sprintf("%.0f", mask_vc_dem)`%).

Interaction models formally test whether concern moderates the partisan gap (see @sec-interaction-models in the Appendix).
For contacts, the Republican × Very Concerned interaction is statistically significant and negative (β = `r sprintf("%.2f", contact_int_concern)`, p = `r sprintf("%.2f", contact_int_concern_p)`), indicating that concern does modestly narrow the partisan contact gap.
For masks, the interaction is large and significant (β = `r sprintf("%.1f", mask_int_concern)`, p \< 0.001).
The mechanism is clear: among the very concerned, Republicans wore masks during `r sprintf("%.0f", mask_veryconcerned_rep)`% of contacts, nearly identical to Democrats (`r sprintf("%.0f", mask_veryconcerned_dem)`%).
But this convergence reflects Republicans responding more sharply to concern: Republicans increased mask usage by `r sprintf("%.0f", mask_slope_rep)` percentage points from "not at all concerned" to "very concerned," compared to `r sprintf("%.0f", mask_slope_dem)` percentage points for Democrats.
Higher perceived severity narrows the mask gap primarily because Republicans increase their masking at a steeper rate.
In sum, perceived severity mitigates partisan polarization for both behaviors, though the effect is substantially larger for mask usage than for social contacts.
These results weigh against the "information and perception" explanation as a complete account: if partisan behavioral differences were simply downstream of different risk perceptions, they should disappear among partisans who express the same level of concern.
They do not — particularly for contacts, where the gap persists at every concern level.
Perception matters, but it is insufficient on its own to explain partisan behavioral divergence.

### How Polarization Evolved Over Time

The cross-sectional analysis above asks whether stated concern explains the partisan gap at a given moment — it does not.
But a complementary question is whether *changes* in objective pandemic severity drove *changes* in the gap over time: did the gap narrow when conditions worsened, as threat override partisan identity, and widen when conditions improved?
To examine this, we track how partisan behavioral gaps evolved across the six survey waves and connect those patterns to local mortality rates partisans would have observed in real time.

@fig-behaviors-over-time displays health behavior levels by party across survey waves (left column) and the corresponding partisan gaps (right column).

```{r}
#| label: fig-behaviors-over-time
#| fig-cap: "Health Behaviors and Partisan Gaps Over Time. Left column shows behavior levels by party; right column shows the absolute partisan gap. For contacts (top row), the gap widened through September 2020 before narrowing during the winter surge. For masks (bottom row), the gap grew steadily over time, suggesting increasing polarization."
#| fig-height: 6
#| fig-width: 10

library(patchwork)
library(cowplot)

# Party colors
party_colors <- c("Democrat" = "#2166AC", "Republican" = "#B2182B", "Independent" = "#6db753")

# Load wave-level data
behaviors_by_wave <- read_csv("../out/tables/behaviors_by_wave.csv", show_col_types = FALSE) %>%
  filter(political_party %in% c("Democrat", "Republican")) %>%
  mutate(
    political_party = factor(political_party, levels = c("Democrat", "Republican")),
    wave_label = factor(wave, labels = c("W1\nApr'20", "W2\nJun'20", "W3\nSep'20",
                                          "W4\nDec'20", "W5\nFeb'21", "W6\nMay'21"))
  )

gaps_by_wave <- read_csv("../out/tables/gaps_by_wave.csv", show_col_types = FALSE) %>%
  mutate(
    wave_label = factor(wave, labels = c("W1\nApr'20", "W2\nJun'20", "W3\nSep'20",
                                          "W4\nDec'20", "W5\nFeb'21", "W6\nMay'21"))
  )

# Panel A: Contacts by party
p_contacts <- ggplot(behaviors_by_wave, aes(x = wave_label, y = total_contacts,
                                             color = political_party, group = political_party)) +
  geom_line(linewidth = 1.2) +
  geom_point(size = 3) +
  scale_color_manual(values = party_colors) +
  labs(x = NULL, y = "Daily Contacts", title = "A. Contacts by Party", color = NULL) +
  theme_cowplot() +
  theme(plot.title = element_text(face = "bold", size = 10),
        axis.text.x = element_text(size = 9),
        legend.position = "none")

# Panel B: Contact gap (absolute)
p_contact_gap <- ggplot(gaps_by_wave, aes(x = wave_label, y = abs(contacts_gap))) +
  geom_col(fill = "#7570b3", alpha = 0.8, color = "black", width = 0.6) +
  geom_text(aes(label = sprintf("%.1f", abs(contacts_gap))), vjust = -0.5, size = 4) +
  labs(x = NULL, y = "Partisan Gap (contacts)", title = "B. Contact Gap", color = NULL) +
  theme_cowplot() +
  theme(plot.title = element_text(face = "bold", size = 10),
        axis.text.x = element_text(size = 9)) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.15)))

# Panel C: Masks by party
p_masks <- ggplot(behaviors_by_wave, aes(x = wave_label, y = mask_usage,
                                          color = political_party, group = political_party)) +
  geom_line(linewidth = 1.2) +
  geom_point(size = 3) +
  scale_color_manual(values = party_colors) +
  labs(x = NULL, y = "Mask Usage (%)", title = "C. Mask Usage by Party", color = NULL) +
  theme_cowplot() +
  theme(plot.title = element_text(face = "bold", size = 10),
        axis.text.x = element_text(size = 9),
        legend.position = "bottom")

# Panel D: Mask gap (absolute)
p_mask_gap <- ggplot(gaps_by_wave, aes(x = wave_label, y = abs(mask_gap))) +
  geom_col(fill = "#1b9e77", alpha = 0.8, color = "black", width = 0.6) +
  geom_text(aes(label = sprintf("%.0f", abs(mask_gap))), vjust = -0.5, size = 4) +
  labs(x = NULL, y = "Partisan Gap (pp)", title = "D. Mask Gap", color = NULL) +
  theme_cowplot() +
  theme(plot.title = element_text(face = "bold", size = 10),
        axis.text.x = element_text(size = 9)) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.15)))

# Combine 2x2
(p_contacts + p_contact_gap) / (p_masks + p_mask_gap) +
  plot_layout(guides = "collect") & theme(legend.position = "bottom")
```

The contact patterns reveal notable variation over time (@fig-behaviors-over-time, top row).
In the previous section, we documented a substantial overall partisan gap in contacts.
But this aggregate difference masks important temporal dynamics.
Two periods showed relatively small contact gaps: early in the pandemic (April–June 2020, gaps of 0.4–0.7 contacts) and during the winter 2020–21 surge (December 2020–February 2021, gaps of 0.3–0.6 contacts).
Between these periods, polarization spiked: the contact gap reached 3.5 contacts in September 2020, nearly ten times larger than during the winter surge.
By May 2021, the gap had widened again to 1.6 contacts.

Mask usage tells a different story (@fig-behaviors-over-time, bottom row).
Rather than narrowing during high-severity periods, the partisan mask gap grew steadily wider over the course of the pandemic.
The gap started at 6 percentage points in April 2020, dipped slightly in September, then climbed to 11 points by December and 15 points by February 2021, where it remained through May.
Unlike contacts, mask polarization showed no convergence during the winter surge; if anything, it accelerated.
This divergence suggests that mask-wearing became an increasingly entrenched marker of partisan identity over time.

What explains these divergent patterns?
One possibility is that partisan gaps in behavior track pandemic conditions as they were reported at the time; specifically, mortality rates that partisans would have observed in real time.[^1]
The timing is suggestive: the periods of smallest contact gaps (early pandemic and winter surge) corresponded to periods of high or rising mortality, while the largest gap (September 2020) occurred when mortality was relatively low (@fig-partisan-gaps).
This pattern suggests Republicans may be more responsive to changes in mortality rates, reducing contacts when deaths rise but also increasing contacts more quickly as rates decline.

[^1]: We focus on mortality rather than incidence here because mortality showed stronger associations with contact behavior.
    See @sec-incidence-vs-mortality for a detailed comparison of incidence and mortality effects.

```{r}
#| label: load-wave-data-perceptions
#| include: false

# Load wave-level data for plotting
gaps_by_wave <- read_csv("../out/tables/gaps_by_wave.csv", show_col_types = FALSE) %>%
  mutate(
    wave_label = case_when(
      wave == 1 ~ "Apr '20",
      wave == 2 ~ "Jun '20",
      wave == 3 ~ "Sep '20",
      wave == 4 ~ "Dec '20",
      wave == 5 ~ "Feb '21",
      wave == 6 ~ "May '21"
    ),
    wave_label = factor(wave_label, levels = c("Apr '20", "Jun '20", "Sep '20", "Dec '20", "Feb '21", "May '21"))
  )
```

```{r}
#| label: fig-partisan-gaps
#| fig-cap: "Partisan Gaps in Health Behaviors Across Survey Waves with Mortality Context. Lines show the Republican-Democrat difference in daily contacts (Panel A) and the Democrat-Republican difference in mask usage (Panel B). Gray bars in background show local mortality rates (deaths per 100,000) during each survey wave. The contact gap was smallest when mortality was highest (Feb '21), while the mask gap grew steadily regardless of mortality."
#| fig-height: 5
#| fig-width: 10

library(patchwork)
library(cowplot)

# Calculate scaling factors for dual axis using mortality
max_mort <- max(gaps_by_wave$mean_mort_rate)
max_contact_gap <- max(gaps_by_wave$contacts_gap)
max_mask_gap <- max(gaps_by_wave$mask_gap)

# Scale factors to make mortality fit nicely in background
contact_scale <- max_contact_gap / max_mort * 1.3
mask_scale <- max_mask_gap / max_mort * 1.3

# Contact gap plot with mortality overlay
p_contact_gap <- ggplot(gaps_by_wave, aes(x = wave_label)) +
  # Background: mortality (scaled)
  geom_col(aes(y = mean_mort_rate * contact_scale),
           fill = "gray70", alpha = 0.5, width = 0.7) +
  # Foreground: partisan gap as line + points
  geom_line(aes(y = contacts_gap, group = 1), color = "#7570b3", linewidth = 1.2) +
  geom_point(aes(y = contacts_gap), color = "#7570b3", size = 4) +
  # Gap value labels
  geom_text(aes(y = contacts_gap, label = sprintf("%.1f", contacts_gap)),
            vjust = -1.2, size = 4, fontface = "bold") +
  geom_hline(yintercept = 0, linetype = "dashed", color = "gray50") +
  scale_y_continuous(
    name = "R-D Gap (contacts/day)",
    sec.axis = sec_axis(~ . / contact_scale, name = "Mortality (per 100k)"),
    expand = expansion(mult = c(0.05, 0.2))
  ) +
  labs(
    x = NULL,
    title = "A. Contact Gap (Republican − Democrat)"
  ) +
  theme_cowplot() +
  theme(
    plot.title = element_text(face = "bold", size = 11),
    axis.text.x = element_text(angle = 45, hjust = 1),
    axis.title.y.right = element_text(color = "gray50")
  )

# Mask gap plot with mortality overlay
p_mask_gap <- ggplot(gaps_by_wave, aes(x = wave_label)) +
  # Background: mortality (scaled)
  geom_col(aes(y = mean_mort_rate * mask_scale),
           fill = "gray70", alpha = 0.5, width = 0.7) +
  # Foreground: partisan gap as line + points
  geom_line(aes(y = mask_gap, group = 1), color = "#1b9e77", linewidth = 1.2) +
  geom_point(aes(y = mask_gap), color = "#1b9e77", size = 4) +
  # Gap value labels
  geom_text(aes(y = mask_gap, label = sprintf("%.1f", mask_gap)),
            vjust = -1.2, size = 4, fontface = "bold") +
  geom_hline(yintercept = 0, linetype = "dashed", color = "gray50") +
  scale_y_continuous(
    name = "D-R Gap (percentage points)",
    sec.axis = sec_axis(~ . / mask_scale, name = "Mortality (per 100k)"),
    expand = expansion(mult = c(0.05, 0.2))
  ) +
  labs(
    x = NULL,
    title = "B. Mask Gap (Democrat − Republican)"
  ) +
  theme_cowplot() +
  theme(
    plot.title = element_text(face = "bold", size = 11),
    axis.text.x = element_text(angle = 45, hjust = 1),
    axis.title.y.right = element_text(color = "gray50")
  )

# Combine
p_contact_gap + p_mask_gap +
  plot_annotation(caption = "Note: Gray background bars show local mortality rates during each survey wave (right axis).")
```

Regression models confirm this pattern.
An interaction between mortality rates and party identification shows that Republicans are indeed more responsive to mortality than Democrats: a one standard deviation increase in local mortality is associated with a reduction of `r sprintf("%.2f", abs(contacts_mort_effect_rep))` contacts for Republicans, compared to `r sprintf("%.2f", abs(contacts_mort_effect_dem))` for Democrats (@fig-mortality-response; see @sec-pandemic-response-detail for detailed coefficients).
Both groups reduce contacts when mortality rises, but Republicans adjust more sharply, consistent with the descriptive pattern of larger swings in Republican contact behavior across the pandemic.

```{r}
#| label: fig-mortality-response
#| fig-cap: "Contact Behavior Response to Mortality. Panel A shows predicted mean daily contacts by party at baseline (average mortality) and when mortality is one standard deviation higher. Panel B shows the Republican-Democrat gap in contacts at baseline and at higher mortality. All partisan groups reduce contacts when mortality rises, narrowing the partisan gap slightly."
#| fig-height: 5
#| fig-width: 10

library(patchwork)
library(cowplot)

# Party colors
party_colors <- c("Democrat" = "#326ba4", "Independent" = "#6db753", "Republican" = "#c50000")

# Calculate predicted means at baseline and +1 SD mortality
# Baseline = raw weighted means; +1 SD = baseline + mortality coefficient
pred_contacts <- tibble(
  party = rep(c("Republican", "Independent", "Democrat"), 2),
  condition = rep(c("Baseline", "+1 SD Mortality"), each = 3),
  contacts = c(
    # Baseline (raw means)
    contacts_rep, contacts_ind, contacts_dem,
    # +1 SD mortality (baseline + effect)
    contacts_rep + contacts_mort_effect_rep,
    contacts_ind + contacts_mort_effect_ind,
    contacts_dem + contacts_mort_effect_dem
  )
) %>%
  mutate(
    party = factor(party, levels = c("Republican", "Independent", "Democrat")),
    condition = factor(condition, levels = c("Baseline", "+1 SD Mortality"))
  )

# Calculate gaps
gap_baseline <- contacts_rep - contacts_dem
gap_high_mort <- (contacts_rep + contacts_mort_effect_rep) - (contacts_dem + contacts_mort_effect_dem)

gap_data <- tibble(
  condition = factor(c("Baseline", "+1 SD Mortality"), levels = c("Baseline", "+1 SD Mortality")),
  gap = c(gap_baseline, gap_high_mort)
)

# Panel A: Predicted contacts by party
p_contacts <- ggplot(pred_contacts, aes(x = condition, y = contacts, fill = party)) +
  geom_col(position = position_dodge(width = 0.8), width = 0.7, alpha = 0.8, color = "black") +
  geom_text(aes(label = sprintf("%.1f", contacts), group = party),
            position = position_dodge(width = 0.8), vjust = -0.5, size = 3.5, fontface = "bold") +
  scale_fill_manual(values = party_colors) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.15))) +
  labs(x = NULL, y = "Mean Daily Contacts", title = "A. Predicted Contacts by Party", fill = "Party") +
  theme_cowplot() +
  theme(
    plot.title = element_text(face = "bold", size = 14),
    legend.position = "bottom"
  )

# Panel B: R-D Gap
p_gap <- ggplot(gap_data, aes(x = condition, y = gap)) +
  geom_col(width = 0.6, fill = "#7570b3", alpha = 0.8, color = "black") +
  geom_text(aes(label = sprintf("%.2f", gap)), vjust = -0.5, size = 4, fontface = "bold") +
  geom_hline(yintercept = 0, linetype = "dashed", color = "gray50") +
  scale_y_continuous(expand = expansion(mult = c(0.05, 0.15))) +
  labs(x = NULL, y = "R-D Gap (contacts)", title = "B. Republican-Democrat Gap") +
  theme_cowplot() +
  theme(
    plot.title = element_text(face = "bold", size = 14)
  )

# Combine
p_contacts + p_gap + plot_layout(widths = c(1.5, 1))
```

### When Attitudes and Behavior Diverge

Yet even as behavior responded to mortality, *stated perceptions* told a different story.
@fig-perception-local-mortality in @sec-perception-vs-local displays each party's concern levels alongside the local mortality rates their respondents experienced.
Early in the pandemic, perception and mortality moved together: in April 2020, when mortality was high, the partisan gap in concern was at its smallest, and Republicans were more similar to Democrats in their likelihood of reporting being "very concerned."
But this alignment was short-lived.
As the pandemic became politicized, stated concern decoupled from both mortality rates and behavior.
The most striking evidence comes from February 2021: mortality reached its peak and the partisan gap in *behavior* narrowed to its smallest point in the study, yet Republican *stated concern* dropped to its lowest (37%; see @fig-perception-levels in @sec-perception-appendix).
Republicans reduced their contacts when deaths surged, but simultaneously reported *less* worry about the pandemic.
This disconnect suggests that stated perceptions became performative over time, reflecting what partisans felt they *should* say rather than the risk assessments actually guiding their behavior.
Stated concern, in other words, increasingly functioned as a marker of partisan identity — a signal of group allegiance — even as underlying behavior continued to respond to objective threat.
The implication is significant: cross-sectional analyses linking stated concern to behavior may partly capture partisan performance rather than genuine risk assessment, and partisan gaps in *attitudes* may overstate the true divergence in the beliefs actually guiding behavior.
For additional analysis of behavior among the "very concerned" over time, including how the contact and mask gaps among this subgroup evolved differently, see @fig-behavior-very-concerned in @sec-behavior-by-concern.

Taken together, these temporal patterns offer a mixed verdict on the "objective risk overrides identity" explanation.
For contacts, the theory finds partial support: when mortality was highest, partisan gaps narrowed, suggesting that concrete, immediate threat can temporarily constrain partisan divergence.
But the theory fails for mask-wearing, where the gap widened steadily regardless of pandemic severity.
Objective risk can moderate partisan polarization for private, high-stakes behaviors — but it does not override identity for visible, low-cost behaviors that have become markers of group membership.
The theory is not eliminated outright, but it cannot account for the full pattern.

## Does Local Context Mitigate Partisan Behavioral Polarization? {#sec-space}

```{r}
#| label: load-region-data
#| include: false

# Load regional data - using adjusted estimates (emmeans with wave + mortality controls)
# These estimates hold wave and mortality constant, removing confounding from
# differences in survey timing and local pandemic severity across regions
behaviors_by_region <- read_csv("../out/tables/adj_behaviors_by_region.csv", show_col_types = FALSE) %>%
  mutate(
    political_party = factor(political_party, levels = c("Republican", "Independent", "Democrat")),
    census_region = factor(census_region, levels = c("Northeast", "Midwest", "South", "West"))
  )

gaps_by_region <- read_csv("../out/tables/adj_gaps_by_region.csv", show_col_types = FALSE) %>%
  mutate(census_region = factor(census_region, levels = c("Northeast", "Midwest", "South", "West")))

# Summary stats for gaps
largest_contact_gap_region <- gaps_by_region %>% slice_max(contacts_gap_rd, n = 1)
smallest_contact_gap_region <- gaps_by_region %>% slice_min(contacts_gap_rd, n = 1)
largest_mask_gap_region <- gaps_by_region %>% slice_max(mask_gap_dr, n = 1)
smallest_mask_gap_region <- gaps_by_region %>% slice_min(mask_gap_dr, n = 1)

# Get behavior levels by party and region for text
dem_contacts <- behaviors_by_region %>% filter(political_party == "Democrat") %>% select(census_region, contacts_adj_mean)
rep_contacts <- behaviors_by_region %>% filter(political_party == "Republican") %>% select(census_region, contacts_adj_mean)
ind_contacts <- behaviors_by_region %>% filter(political_party == "Independent") %>% select(census_region, contacts_adj_mean)
dem_masks <- behaviors_by_region %>% filter(political_party == "Democrat") %>% select(census_region, mask_adj_mean)
rep_masks <- behaviors_by_region %>% filter(political_party == "Republican") %>% select(census_region, mask_adj_mean)
ind_masks <- behaviors_by_region %>% filter(political_party == "Independent") %>% select(census_region, mask_adj_mean)

# Calculate variation (range) by party for contacts
dem_contacts_range <- max(dem_contacts$contacts_adj_mean) - min(dem_contacts$contacts_adj_mean)
rep_contacts_range <- max(rep_contacts$contacts_adj_mean) - min(rep_contacts$contacts_adj_mean)
ind_contacts_range <- max(ind_contacts$contacts_adj_mean) - min(ind_contacts$contacts_adj_mean)

# Calculate variation (range) by party for masks
dem_masks_range <- max(dem_masks$mask_adj_mean) - min(dem_masks$mask_adj_mean)
rep_masks_range <- max(rep_masks$mask_adj_mean) - min(rep_masks$mask_adj_mean)
ind_masks_range <- max(ind_masks$mask_adj_mean) - min(ind_masks$mask_adj_mean)

# Regional values by party
rep_south_contacts <- rep_contacts %>% filter(census_region == "South") %>% pull(contacts_adj_mean)
rep_midwest_contacts <- rep_contacts %>% filter(census_region == "Midwest") %>% pull(contacts_adj_mean)
rep_south_masks <- rep_masks %>% filter(census_region == "South") %>% pull(mask_adj_mean)
dem_south_contacts <- dem_contacts %>% filter(census_region == "South") %>% pull(contacts_adj_mean)
dem_south_masks <- dem_masks %>% filter(census_region == "South") %>% pull(mask_adj_mean)
dem_midwest_contacts <- dem_contacts %>% filter(census_region == "Midwest") %>% pull(contacts_adj_mean)
dem_other_contacts <- dem_contacts %>% filter(census_region != "South") %>% summarise(mean = mean(contacts_adj_mean)) %>% pull(mean)
ind_south_contacts <- ind_contacts %>% filter(census_region == "South") %>% pull(contacts_adj_mean)
ind_midwest_contacts <- ind_contacts %>% filter(census_region == "Midwest") %>% pull(contacts_adj_mean)
ind_south_masks <- ind_masks %>% filter(census_region == "South") %>% pull(mask_adj_mean)

# Create region-to-state mapping (used by multiple figures)
library(usmap)
library(sf)

region_map <- tibble(
  state = c(
    "Connecticut", "Maine", "Massachusetts", "New Hampshire", "New Jersey",
    "New York", "Pennsylvania", "Rhode Island", "Vermont",
    "Illinois", "Indiana", "Iowa", "Kansas", "Michigan", "Minnesota",
    "Missouri", "Nebraska", "North Dakota", "Ohio", "South Dakota", "Wisconsin",
    "Alabama", "Arkansas", "Delaware", "District of Columbia", "Florida",
    "Georgia", "Kentucky", "Louisiana", "Maryland", "Mississippi",
    "North Carolina", "Oklahoma", "South Carolina", "Tennessee", "Texas",
    "Virginia", "West Virginia",
    "Alaska", "Arizona", "California", "Colorado", "Hawaii", "Idaho",
    "Montana", "Nevada", "New Mexico", "Oregon", "Utah", "Washington", "Wyoming"
  ),
  census_region = c(
    rep("Northeast", 9), rep("Midwest", 12), rep("South", 17), rep("West", 13)
  )
)

# Get state polygons and create region boundaries (used by multiple figures)
us_states_sf <- us_map(regions = "states") %>%
  left_join(region_map, by = c("full" = "state"))

region_boundaries <- us_states_sf %>%
  group_by(census_region) %>%
  summarise(geometry = st_union(geom), .groups = "drop")
```

The preceding analysis focused on how pandemic severity, whether perceived or objective, shaped partisan behavioral differences over time.
We now turn to whether *where* people live affects partisan polarization.
If partisan gaps reflect local policy environments, regional culture, or adaptation to local norms, they should narrow or disappear within shared geographic contexts.

### Geographic and Policy Variation

@fig-region-gap-map displays Republican-Democrat gaps in contacts and mask usage across census regions, adjusted for survey wave and local mortality rates (behavior levels by party and region are shown in @fig-region-levels-map in @sec-spatial-variation).

```{r}
#| label: fig-region-levels-map-setup
#| include: false

library(patchwork)

# Reshape behaviors data for mapping
rep_data <- behaviors_by_region %>%
  filter(political_party == "Republican") %>%
  select(census_region, contacts_adj_mean, mask_adj_mean) %>%
  rename(contacts_mean = contacts_adj_mean, mask_mean = mask_adj_mean)

ind_data <- behaviors_by_region %>%
  filter(political_party == "Independent") %>%
  select(census_region, contacts_adj_mean, mask_adj_mean) %>%
  rename(contacts_mean = contacts_adj_mean, mask_mean = mask_adj_mean)

dem_data <- behaviors_by_region %>%
  filter(political_party == "Democrat") %>%
  select(census_region, contacts_adj_mean, mask_adj_mean) %>%
  rename(contacts_mean = contacts_adj_mean, mask_mean = mask_adj_mean)

rep_for_map <- region_map %>% left_join(rep_data, by = "census_region")
ind_for_map <- region_map %>% left_join(ind_data, by = "census_region")
dem_for_map <- region_map %>% left_join(dem_data, by = "census_region")

# Set common scales across all three parties
contacts_range <- range(c(rep_data$contacts_mean, ind_data$contacts_mean, dem_data$contacts_mean), na.rm = TRUE)
mask_range <- range(c(rep_data$mask_mean, ind_data$mask_mean, dem_data$mask_mean), na.rm = TRUE)

# Republican contacts
p_rep_contacts <- plot_usmap(data = rep_for_map, values = "contacts_mean", color = "white", linewidth = 0.2) +
  geom_sf(data = region_boundaries, fill = NA, color = "black", linewidth = 0.6, inherit.aes = FALSE) +
  scale_fill_gradient(low = "#fee0d2", high = "#cb181d", name = "Contacts", limits = contacts_range, na.value = "gray90") +
  labs(title = "A. Republican Contacts") +
  theme(plot.title = element_text(face = "bold", size = 11, hjust = 0.5), legend.position = "right")

# Republican masks
p_rep_masks <- plot_usmap(data = rep_for_map, values = "mask_mean", color = "white", linewidth = 0.2) +
  geom_sf(data = region_boundaries, fill = NA, color = "black", linewidth = 0.6, inherit.aes = FALSE) +
  scale_fill_gradient(low = "#fee0d2", high = "#cb181d", name = "Mask %", limits = mask_range, na.value = "gray90") +
  labs(title = "B. Republican Mask Usage") +
  theme(plot.title = element_text(face = "bold", size = 11, hjust = 0.5), legend.position = "right")

# Independent contacts
p_ind_contacts <- plot_usmap(data = ind_for_map, values = "contacts_mean", color = "white", linewidth = 0.2) +
  geom_sf(data = region_boundaries, fill = NA, color = "black", linewidth = 0.6, inherit.aes = FALSE) +
  scale_fill_gradient(low = "#e5f5e0", high = "#31a354", name = "Contacts", limits = contacts_range, na.value = "gray90") +
  labs(title = "C. Independent Contacts") +
  theme(plot.title = element_text(face = "bold", size = 11, hjust = 0.5), legend.position = "right")

# Independent masks
p_ind_masks <- plot_usmap(data = ind_for_map, values = "mask_mean", color = "white", linewidth = 0.2) +
  geom_sf(data = region_boundaries, fill = NA, color = "black", linewidth = 0.6, inherit.aes = FALSE) +
  scale_fill_gradient(low = "#e5f5e0", high = "#31a354", name = "Mask %", limits = mask_range, na.value = "gray90") +
  labs(title = "D. Independent Mask Usage") +
  theme(plot.title = element_text(face = "bold", size = 11, hjust = 0.5), legend.position = "right")

# Democrat contacts
p_dem_contacts <- plot_usmap(data = dem_for_map, values = "contacts_mean", color = "white", linewidth = 0.2) +
  geom_sf(data = region_boundaries, fill = NA, color = "black", linewidth = 0.6, inherit.aes = FALSE) +
  scale_fill_gradient(low = "#deebf7", high = "#2171b5", name = "Contacts", limits = contacts_range, na.value = "gray90") +
  labs(title = "E. Democrat Contacts") +
  theme(plot.title = element_text(face = "bold", size = 11, hjust = 0.5), legend.position = "right")

# Democrat masks
p_dem_masks <- plot_usmap(data = dem_for_map, values = "mask_mean", color = "white", linewidth = 0.2) +
  geom_sf(data = region_boundaries, fill = NA, color = "black", linewidth = 0.6, inherit.aes = FALSE) +
  scale_fill_gradient(low = "#deebf7", high = "#2171b5", name = "Mask %", limits = mask_range, na.value = "gray90") +
  labs(title = "F. Democrat Mask Usage") +
  theme(plot.title = element_text(face = "bold", size = 11, hjust = 0.5), legend.position = "right")

# Combine in 3x2 grid: rows = party, columns = behavior
(p_rep_contacts | p_rep_masks) / (p_ind_contacts | p_ind_masks) / (p_dem_contacts | p_dem_masks)
```

Partisan gaps persist across every region: Republicans reported more daily contacts and lower mask usage than Democrats regardless of where they lived.
The contact gap was largest in the `r largest_contact_gap_region$census_region` (`r sprintf("%.1f", largest_contact_gap_region$contacts_gap_rd)` contacts) and smallest in the `r smallest_contact_gap_region$census_region` (`r sprintf("%.1f", smallest_contact_gap_region$contacts_gap_rd)` contacts); the mask gap was similarly largest in the `r largest_mask_gap_region$census_region` (`r sprintf("%.1f", largest_mask_gap_region$mask_gap_dr)` pp) and smallest in the `r smallest_mask_gap_region$census_region` (`r sprintf("%.1f", smallest_mask_gap_region$mask_gap_dr)` pp).
Notably, the smaller gap in the South is driven primarily by Democrats reporting more contacts (`r sprintf("%.1f", dem_south_contacts)`) than Democrats elsewhere (`r sprintf("%.1f", dem_other_contacts)` on average) — consistent with Democrats in a heavily Republican region adjusting toward local norms — rather than by Southern Republicans becoming more cautious.[^geo]

[^geo]: See @sec-spatial-variation for additional geographic breakdowns including vaccination, perception, and finer-grained division and state-level maps.

```{r}
#| label: fig-region-gap-map
#| fig-cap: "Geographic Distribution of Partisan Gaps by Census Region (Adjusted). Panel A shows the absolute partisan gap in daily contacts; Panel B shows the absolute partisan gap in mask usage. Estimates are adjusted for survey wave and local mortality rates. Darker shading indicates larger gaps. Black borders indicate census region boundaries."
#| fig-height: 8
#| fig-width: 8

library(patchwork)

# Merge with gaps data
region_gaps_for_map <- region_map %>%
  left_join(gaps_by_region, by = "census_region")

# Contact gap map - purple gradient (all regions have R > D)
p_contact_map <- plot_usmap(data = region_gaps_for_map, values = "contacts_gap_rd", color = "white", linewidth = 0.2) +
  geom_sf(data = region_boundaries, fill = NA, color = "black", linewidth = 0.8, inherit.aes = FALSE) +
  scale_fill_gradient(
    low = "#f2e6ff", high = "#5c2d91",
    name = "Partisan Gap\n(contacts)",
    na.value = "gray90"
  ) +
  labs(title = "A. Contact Gap") +
  theme(
    plot.title = element_text(face = "bold", size = 12, hjust = 0.5),
    legend.position = "right"
  )

# Mask gap map - gold gradient (all regions have D > R)
p_mask_map <- plot_usmap(data = region_gaps_for_map, values = "mask_gap_dr", color = "white", linewidth = 0.2) +
  geom_sf(data = region_boundaries, fill = NA, color = "black", linewidth = 0.8, inherit.aes = FALSE) +
  scale_fill_gradient(
    low = "#fff7e6", high = "#b8860b",
    name = "Partisan Gap\n(mask usage)",
    na.value = "gray90"
  ) +
  labs(title = "B. Mask Gap") +
  theme(
    plot.title = element_text(face = "bold", size = 12, hjust = 0.5),
    legend.position = "right"
  )

# Stack vertically
p_contact_map / p_mask_map
```

These regional patterns are robust to controls for local context and demographic composition.
Partisan gaps persisted within each policy environment (@fig-policy-behaviors in @sec-policy-behaviors) and remained stable after adjusting for urbanicity and demographics: contact gaps ranged from 0.7 contacts in the South to 1.5–1.7 contacts elsewhere with minimal change across model specifications; mask gaps shrank modestly but persisted everywhere, with the West retaining a 12-point gap in the fully adjusted model (see @fig-region-gap-context-adjustment in @sec-context-demographics-robustness).

These findings eliminate two competing explanations.
The "it's just policy" theory predicts that partisan gaps should disappear among people living under the same mandate regimes — they do not.
The "context and culture determine behavior" theory predicts that regional culture should override partisan identity, such that partisans within the South or within metropolitan areas should behave similarly regardless of party — they do not.
Partisan gaps persist within every policy environment, every region, and every level of urbanicity, ruling out policy and local culture as sufficient explanations.
Supplementary analyses of partisans living in opposing-party congressional districts tell a consistent story: while partisans do adjust modestly toward local norms, the partisan gap persists in both same-party and cross-party contexts (see @sec-cd-context-robustness).

Taken together, the full set of results points to a single surviving explanation: partisan identity itself drives behavioral polarization.
The gaps cannot be reduced to demographics, risk perception, objective severity, policy environments, or regional culture.
Each of these factors matters at the margins — perception narrows the mask gap among the very concerned, mortality temporarily compresses the contact gap, and regional context shifts absolute levels — but none eliminates the fundamental partisan divide.
What remains is "portable polarization": partisan identity travels with individuals, shaping their health behaviors regardless of where they live, what policies govern them, or what risks surround them.
