# Outtakes

This file holds cut text preserved for potential reuse.

---

## Introduction: Competing theories (paragraph-by-paragraph version)

*Cut from introduction in favor of compressed single paragraph. Consider using this logic at the top of the results or as an analytical framework subsection after methods.*

Several theoretical perspectives generate distinct predictions about how partisan gaps should vary across contexts.
If partisan differences are primarily artifacts of policy environments, gaps should disappear when we compare partisans living under similar policies—within strict mandate areas, within lenient areas, and within areas with no mandates.
If partisan differences reflect demographic composition, controlling for age, education, urbanicity, and other covariates should eliminate the gaps.
If local social conformity shapes behavior, we should observe individuals shifting toward the norms of their local community regardless of their own partisan identity—Democrats in Republican areas behaving more like Republicans, and vice versa.
If cross-cutting social contexts reduce polarization, partisan gaps should be smaller when partisans live among members of the opposing party than when they live among co-partisans.

A key question is whether local context fully determines behavior or whether partisan identity exerts an independent effect.
If context dominates, partisans in the same environment should behave similarly—a Democrat in a Republican-majority area should be nearly indistinguishable from their Republican neighbors.
If partisan identity dominates, gaps should persist even within shared contexts: Democrats and Republicans living in the same types of areas should still behave differently from one another, even if both groups shift their behavior in response to local conditions.
The persistence of partisan gaps across multiple contextual dimensions—partisan composition, urbanicity, region, and policy environment—would provide strong evidence that partisan identity drives health behavior independent of local culture.
It would also suggest that behavioral polarization is "portable": people carry their partisan patterns with them regardless of where they live, implying that geographic mixing alone will not reduce polarization.

An alternative explanation emphasizes information rather than identity.
Republicans and Democrats may have been exposed to different information about the pandemic through divergent media ecosystems and elite messaging \[CITE\].
If so, partisan differences in behavior could simply reflect partisan differences in perceived severity: Republicans behaved less cautiously because they genuinely believed the pandemic was less dangerous, based on the information they received.
This explanation generates specific predictions.
First, we should observe partisan differences in perceptions of pandemic severity—which prior research confirms \[CITE\].
Second, and critically, perceptions should strongly predict behavior: those who perceived the pandemic as more severe should have behaved more cautiously.
If information and perceptions fully mediate the relationship between partisanship and behavior, controlling for perceptions should substantially reduce partisan gaps.
However, if partisan identity shapes behavior through channels beyond information processing—through social conformity, symbolic expression of group membership, or resistance to outgroup norms—behavioral gaps may persist even among partisans with similar perceptions.
A third possibility is that stated perceptions are themselves expressions of partisan identity: partisans may report what their group is "supposed" to believe about pandemic severity, even if their actual behavior reveals different risk assessments.
If Republicans report low concern but behave more cautiously than their stated perceptions would predict, this would suggest that expressed perceptions are partly performative—reflecting what one says as a group member rather than genuine beliefs that drive behavior.
We examine whether perceptions and behaviors move together, and whether behavioral gaps remain after accounting for perceived severity.

Finally, we examine whether the severity of the pandemic moderates partisan polarization.
If polarization thrives when threats are abstract but narrows when consequences become concrete and personal, partisan gaps should be smaller during periods of high mortality and incidence, when the pandemic's toll was most visible and immediate.
Conversely, gaps may widen during periods of lower severity, when there was more room for motivated reasoning and partisan interpretation of ambiguous risk information.
Such a pattern would suggest that reality can partially override partisan identity when stakes are high enough—but that polarization may return when the immediate threat recedes.

---

## Appendix Outtakes (moved during JHSB revision)

*The following appendix sections were removed from the main appendix because they are not directly cross-referenced from the main text. Preserved here for potential reuse.*

<!-- === Appendix C: Stay-in-Place Orders Over Time === -->

## APPENDIX C. Stay-in-Place Orders Over Time {#sec-sip-appendix}

@fig-sip-over-time shows the percentage of respondents reporting being under a stay-in-place order across the study period. The BICS survey asked respondents whether they were currently under a stay-in-place order in waves 1-5, but dropped this question entirely in wave 6 (May 2021)---the only survey item discontinued during the study---reflecting the widespread lifting of restrictions by that time.

```{r}
#| label: fig-sip-over-time
#| fig-cap: "Percentage Reporting Being Under a Stay-in-Place Order Over Time. Panel A shows all respondents; Panel B shows only those reporting being 'very concerned' about COVID-19. Stay-in-place orders declined substantially over the study period. The question was dropped in wave 6 as most states had lifted formal restrictions by May 2021."
#| fig-height: 4
#| fig-width: 10

library(patchwork)
library(cowplot)

# Load SIP data for all respondents
pct_sip_all <- read_csv("../../out/tables/pct_under_sip_by_wave.csv", show_col_types = FALSE) %>%
  filter(political_party %in% c("Democrat", "Republican")) %>%
  mutate(
    political_party = factor(political_party, levels = c("Democrat", "Republican")),
    wave_label = factor(wave, levels = 1:6, labels = c("W1\nApr'20", "W2\nJun'20", "W3\nSep'20",
                                          "W4\nDec'20", "W5\nFeb'21", "W6\nMay'21"))
  )

# Load SIP data for very concerned only
pct_sip_vc <- read_csv("../../out/tables/pct_under_sip_very_concerned_by_wave.csv", show_col_types = FALSE) %>%
  filter(political_party %in% c("Democrat", "Republican")) %>%
  mutate(
    political_party = factor(political_party, levels = c("Democrat", "Republican")),
    wave_label = factor(wave, levels = 1:6, labels = c("W1\nApr'20", "W2\nJun'20", "W3\nSep'20",
                                          "W4\nDec'20", "W5\nFeb'21", "W6\nMay'21"))
  )

# Panel A: All respondents
p_all <- ggplot(pct_sip_all, aes(x = wave_label, y = pct_under_sip, color = political_party, group = political_party)) +
  geom_line(linewidth = 1.2) +
  geom_point(size = 3) +
  scale_x_discrete(drop = FALSE) +
  scale_color_manual(values = c("Democrat" = "#2166AC", "Republican" = "#B2182B")) +
  labs(x = NULL, y = "% Under Stay-in-Place", title = "A. All Respondents", color = NULL) +
  theme_cowplot() +
  theme(plot.title = element_text(face = "bold", size = 11), axis.text.x = element_text(size = 8)) +
  annotate("text", x = 6, y = 75, label = "Question\ndropped", size = 2.5, hjust = 0.5, color = "gray40") +
  ylim(60, 100)

# Panel B: Very concerned only
p_vc <- ggplot(pct_sip_vc, aes(x = wave_label, y = pct_under_sip, color = political_party, group = political_party)) +
  geom_line(linewidth = 1.2) +
  geom_point(size = 3) +
  scale_x_discrete(drop = FALSE) +
  scale_color_manual(values = c("Democrat" = "#2166AC", "Republican" = "#B2182B")) +
  labs(x = NULL, y = "% Under Stay-in-Place", title = "B. 'Very Concerned' Only", color = NULL) +
  theme_cowplot() +
  theme(plot.title = element_text(face = "bold", size = 11), axis.text.x = element_text(size = 8)) +
  annotate("text", x = 6, y = 82, label = "Question\ndropped", size = 2.5, hjust = 0.5, color = "gray40") +
  ylim(70, 100)

p_all + p_vc + plot_layout(guides = "collect") & theme(legend.position = "bottom")
```

Among all respondents (Panel A), the percentage reporting being under a stay-in-place order declined from approximately 95% in wave 1 to 70-81% by wave 5, with Republicans showing a larger decline than Democrats (11 percentage point gap by wave 5). This overall pattern is more representative than the "very concerned" subset (Panel B), since the share of Republicans reporting being "very concerned" declined substantially over the study---from 55% in wave 1 to just 26% by wave 6---making the very concerned Republicans an increasingly selective group. The declining trend in stay-in-place orders provides context for the behavioral changes observed in the main text: as restrictions lifted, particularly between waves 4-6, Democrats increased their contacts while maintaining high mask usage---consistent with a strategic shift toward relying on masks and vaccines for protection as they resumed social activities.


<!-- === Appendix E: Adjusted Partisan Behaviors Over Time === -->

## APPENDIX E. Adjusted Partisan Behaviors Over Time {#sec-adjusted-behaviors}

```{r}
#| label: load-adjusted-data
#| include: false

# Load adjusted behaviors data (from emmeans models)
adjusted_behaviors <- read_csv("../../out/tables/adjusted_behaviors_by_wave.csv", show_col_types = FALSE) %>%
  mutate(
    political_party = factor(political_party, levels = c("Republican", "Independent", "Democrat")),
    wave_label = case_when(
      wave == 1 ~ "Apr '20",
      wave == 2 ~ "Jun '20",
      wave == 3 ~ "Sep '20",
      wave == 4 ~ "Dec '20",
      wave == 5 ~ "Feb '21",
      wave == 6 ~ "May '21"
    ),
    wave_label = factor(wave_label, levels = c("Apr '20", "Jun '20", "Sep '20", "Dec '20", "Feb '21", "May '21"))
  )

adjusted_gaps <- read_csv("../../out/tables/adjusted_gaps_by_wave.csv", show_col_types = FALSE) %>%
  mutate(
    wave_label = case_when(
      wave == 1 ~ "Apr '20",
      wave == 2 ~ "Jun '20",
      wave == 3 ~ "Sep '20",
      wave == 4 ~ "Dec '20",
      wave == 5 ~ "Feb '21",
      wave == 6 ~ "May '21"
    ),
    wave_label = factor(wave_label, levels = c("Apr '20", "Jun '20", "Sep '20", "Dec '20", "Feb '21", "May '21"))
  )

# Separate by outcome and filter to just Republicans and Democrats
contacts_adjusted <- adjusted_behaviors %>%
  filter(outcome == "contacts", political_party != "Independent")
masks_adjusted <- adjusted_behaviors %>%
  filter(outcome == "masks", political_party != "Independent")
concern_adjusted <- adjusted_behaviors %>%
  filter(outcome == "concern", political_party != "Independent")

contacts_gaps <- adjusted_gaps %>% filter(outcome == "contacts")
masks_gaps <- adjusted_gaps %>% filter(outcome == "masks")
concern_gaps <- adjusted_gaps %>% filter(outcome == "concern")

# Summary stats for text
max_contact_gap_adj <- contacts_gaps %>% slice_max(gap_rep_dem, n = 1)
min_contact_gap_adj <- contacts_gaps %>% slice_min(gap_rep_dem, n = 1)
max_mask_gap_adj <- masks_gaps %>% slice_max(abs(gap_rep_dem), n = 1)
min_mask_gap_adj <- masks_gaps %>% slice_min(abs(gap_rep_dem), n = 1)
max_concern_gap_adj <- concern_gaps %>% slice_max(abs(gap_rep_dem), n = 1)
min_concern_gap_adj <- concern_gaps %>% slice_min(abs(gap_rep_dem), n = 1)
```

The main text presents descriptive partisan differences in health behaviors over time. To assess whether these patterns persist after accounting for demographic and contextual factors, we estimate regression models with party-by-wave interactions, controlling for age, race/ethnicity, Hispanic origin, gender, education, household size, congressional district partisan composition, and local mortality rates. We then extract estimated marginal means (emmeans) for each party at each wave, averaging over covariates. This approach isolates partisan differences while holding constant the demographic and geographic factors that might otherwise confound comparisons.

@fig-adjusted-behaviors displays the adjusted behavior levels by party across survey waves, and @fig-adjusted-gaps shows the resulting partisan gaps.

```{r}
#| label: fig-adjusted-behaviors
#| fig-cap: "Adjusted Health Behaviors and Perceptions for Republicans and Democrats Across Survey Waves. Points show estimated marginal means from models controlling for age, race/ethnicity, gender, education, household size, congressional district partisanship, and local mortality rates. Error bars indicate 95% confidence intervals."
#| fig-height: 4
#| fig-width: 14

library(cowplot)
library(patchwork)

# Party colors
party_colors <- c("Republican" = "#c50000", "Independent" = "#6db753", "Democrat" = "#326ba4")

# Contacts plot
p_contacts_adj <- ggplot(contacts_adjusted, aes(x = wave_label, y = estimate,
                                                 color = political_party, group = political_party)) +
  geom_line(linewidth = 1.2) +
  geom_point(size = 3) +
  geom_errorbar(aes(ymin = lower.CL, ymax = upper.CL), width = 0.15, alpha = 0.7) +
  scale_color_manual(values = party_colors) +
  labs(x = NULL, y = "Daily Contacts", title = "A. Adjusted Daily Contacts", color = "Party") +
  theme_cowplot() +
  theme(
    plot.title = element_text(face = "bold", size = 11),
    axis.text.x = element_text(angle = 45, hjust = 1),
    legend.position = "bottom"
  )

# Masks plot
p_masks_adj <- ggplot(masks_adjusted, aes(x = wave_label, y = estimate,
                                           color = political_party, group = political_party)) +
  geom_line(linewidth = 1.2) +
  geom_point(size = 3) +
  geom_errorbar(aes(ymin = lower.CL, ymax = upper.CL), width = 0.15, alpha = 0.7) +
  scale_color_manual(values = party_colors) +
  labs(x = NULL, y = "Mask Usage (%)", title = "B. Adjusted Mask Usage", color = "Party") +
  theme_cowplot() +
  theme(
    plot.title = element_text(face = "bold", size = 11),
    axis.text.x = element_text(angle = 45, hjust = 1),
    legend.position = "bottom"
  )

# Concern plot
p_concern_adj <- ggplot(concern_adjusted, aes(x = wave_label, y = estimate,
                                               color = political_party, group = political_party)) +
  geom_line(linewidth = 1.2) +
  geom_point(size = 3) +
  geom_errorbar(aes(ymin = lower.CL, ymax = upper.CL), width = 0.15, alpha = 0.7) +
  scale_color_manual(values = party_colors) +
  labs(x = NULL, y = "Very Concerned (%)", title = "C. Adjusted Concern", color = "Party") +
  theme_cowplot() +
  theme(
    plot.title = element_text(face = "bold", size = 11),
    axis.text.x = element_text(angle = 45, hjust = 1),
    legend.position = "bottom"
  )

p_contacts_adj + p_masks_adj + p_concern_adj + plot_layout(guides = "collect") &
  theme(legend.position = "bottom")
```

```{r}
#| label: fig-adjusted-gaps
#| fig-cap: "Adjusted Partisan Gaps in Health Behaviors and Perceptions Over Time. Bars show the Republican-Democrat difference in estimated marginal means for daily contacts (Panel A), mask usage (Panel B), and stated concern (Panel C), after controlling for covariates. Negative values indicate Democrats are higher. Purple bars indicate statistically significant gaps (p < 0.05); gray bars indicate non-significant gaps."
#| fig-height: 4
#| fig-width: 14

# Calculate standard errors for gaps and determine significance
# SE_gap = sqrt(SE_R^2 + SE_D^2)
calc_gap_significance <- function(behaviors_df, outcome_name) {
  behaviors_df %>%
    filter(outcome == outcome_name) %>%
    select(wave, political_party, estimate, std_error) %>%
    pivot_wider(names_from = political_party, values_from = c(estimate, std_error)) %>%
    mutate(
      gap_rep_dem = estimate_Republican - estimate_Democrat,
      se_gap = sqrt(std_error_Republican^2 + std_error_Democrat^2),
      significant = abs(gap_rep_dem) > 1.96 * se_gap
    ) %>%
    select(wave, gap_rep_dem, se_gap, significant)
}

contacts_sig <- calc_gap_significance(adjusted_behaviors, "contacts") %>%
  mutate(outcome_label = "A. Contact Gap (R - D)")
masks_sig <- calc_gap_significance(adjusted_behaviors, "masks") %>%
  mutate(outcome_label = "B. Mask Gap (R - D)")
concern_sig <- calc_gap_significance(adjusted_behaviors, "concern") %>%
  mutate(outcome_label = "C. Concern Gap (R - D)")

# Combine with wave labels
gaps_long <- bind_rows(contacts_sig, masks_sig, concern_sig) %>%
  mutate(
    wave_label = case_when(
      wave == 1 ~ "Apr '20",
      wave == 2 ~ "Jun '20",
      wave == 3 ~ "Sep '20",
      wave == 4 ~ "Dec '20",
      wave == 5 ~ "Feb '21",
      wave == 6 ~ "May '21"
    ),
    wave_label = factor(wave_label, levels = c("Apr '20", "Jun '20", "Sep '20", "Dec '20", "Feb '21", "May '21"))
  )

ggplot(gaps_long, aes(x = wave_label, y = gap_rep_dem, fill = significant)) +
  geom_col(alpha = 0.85, width = 0.6, color = "black") +
  geom_hline(yintercept = 0, linetype = "dashed", color = "gray50") +
  geom_text(aes(label = sprintf("%.1f", gap_rep_dem),
                y = gap_rep_dem / 2),
            color = "white", size = 3.5, fontface = "bold") +
  scale_fill_manual(values = c("TRUE" = "#7570b3", "FALSE" = "gray60"),
                    labels = c("TRUE" = "p < 0.05", "FALSE" = "n.s."),
                    name = NULL) +
  facet_wrap(~ outcome_label, scales = "free_y", nrow = 1) +
  labs(x = NULL, y = "Adjusted R-D Difference") +
  theme_cowplot() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    strip.text = element_text(face = "bold"),
    strip.background = element_blank(),
    legend.position = "bottom"
  )
```

### Interpretation

After adjusting for covariates, the patterns of partisan differences reveal important nuances. For **contact behavior**, the descriptive data show that partisan gaps were smallest early in the pandemic, with Republicans reporting only 0.4 more contacts than Democrats in April 2020 compared to gaps of 1-3 contacts in later waves. The adjusted analysis goes further: after controlling for demographics, the April 2020 gap is not only small but *statistically indistinguishable from zero* (see @fig-adjusted-gaps, gray bars indicate non-significant differences). In fact, the point estimate is slightly negative, suggesting that if anything, Democrats had marginally more contacts than Republicans when mortality was at its peak. This stands in stark contrast to the large, statistically significant gaps observed in September 2020 and May 2021, periods of relatively lower pandemic intensity. The pattern supports the interpretation that concrete, visible threat can eliminate---not merely reduce---partisan divergence in behavior.

For **mask-wearing**, the adjusted analysis reveals that partisan differences fluctuated rather than monotonically increasing. The Democrat-Republican mask gap (with Democrats wearing masks more) ranged from `r sprintf("%.1f", abs(min_mask_gap_adj$gap_rep_dem))` percentage points in `r min_mask_gap_adj$wave_label` to `r sprintf("%.1f", abs(max_mask_gap_adj$gap_rep_dem))` percentage points in `r max_mask_gap_adj$wave_label`. Notably, the gap was relatively narrow during fall 2020 when mask usage was high across all groups, but widened again in early 2021 as vaccines became available and Republicans began to decrease mask usage more rapidly.

For **stated concern**, the adjusted analysis confirms that the perception gap was smallest at the beginning of the pandemic (when mortality was highest) and widened substantially thereafter. The adjusted Democrat-Republican concern gap ranged from `r sprintf("%.1f", abs(min_concern_gap_adj$gap_rep_dem))` percentage points in `r min_concern_gap_adj$wave_label` to `r sprintf("%.1f", abs(max_concern_gap_adj$gap_rep_dem))` percentage points in `r max_concern_gap_adj$wave_label`. This pattern parallels the contact findings: when objective threat was most severe, partisan gaps in both behavior and perception were smallest. The subsequent widening of the concern gap, even as mortality rates declined, supports the interpretation that politicization drove perceptual divergence.

These adjusted estimates confirm that partisan differences in health behaviors and perceptions were not simply artifacts of demographic composition or geographic sorting. Even after controlling for a comprehensive set of covariates, substantial partisan gaps persisted throughout the pandemic, though their magnitude varied with pandemic conditions---narrowing when mortality was highest and widening as the pandemic became increasingly politicized.


<!-- === Appendix F: Partisan Differences Within Policy Environments (duplicate of sec-policy-behaviors) === -->

## APPENDIX F. Partisan Differences Within Policy Environments {#sec-policy-environment}

```{r}
#| label: load-policy-data
#| include: false

# Load behaviors by policy environment
behaviors_by_policy <- read_csv("../../out/tables/behaviors_by_policy.csv", show_col_types = FALSE) %>%
  mutate(
    political_party = factor(political_party, levels = c("Republican", "Independent", "Democrat")),
    policy = factor(policy, levels = c("No Mandate", "Less Strict", "Strict"))
  )
```

A potential alternative explanation for partisan differences in health behaviors is that Democrats disproportionately reside in jurisdictions with stricter COVID-19 policies, and their behavior simply reflects compliance with local regulations rather than partisan preferences. To address this concern, we examine whether partisan differences persist within the same policy environment.

@fig-contacts-by-policy and @fig-masks-by-policy display mean contact rates and mask usage by party, stratified by county-level mask mandate policy (No Mandate, Less Strict, or Strict). If policy environment fully explained partisan differences, we would expect the gaps between parties to disappear within each policy category.

```{r}
#| label: fig-contacts-by-policy
#| fig-cap: "Daily Contacts by Party Within Policy Environments. Bars show weighted mean daily contacts for each party, faceted by county mask mandate policy. Error bars indicate standard errors."
#| fig-height: 8
#| fig-width: 6

library(cowplot)

# Party colors
party_colors <- c("Republican" = "#c50000", "Independent" = "#6db753", "Democrat" = "#326ba4")

# Contacts by policy - faceted vertically
ggplot(behaviors_by_policy, aes(x = political_party, y = contacts_mean, fill = political_party)) +
  geom_col(width = 0.7, alpha = 0.8, color = "black") +
  geom_errorbar(aes(ymin = contacts_mean - contacts_se, ymax = contacts_mean + contacts_se),
                width = 0.2) +
  geom_text(aes(label = sprintf("%.1f", contacts_mean)), vjust = -1.2, size = 3.5, fontface = "bold") +
  facet_wrap(~ policy, ncol = 1) +
  scale_fill_manual(values = party_colors) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.2))) +
  labs(x = NULL, y = "Daily Contacts") +
  theme_cowplot() +
  theme(
    legend.position = "none",
    axis.text.x = element_text(angle = 45, hjust = 1, size = 11),
    strip.text = element_text(face = "bold", size = 14),
    strip.background = element_blank()
  )
```

```{r}
#| label: fig-masks-by-policy
#| fig-cap: "Mask Usage by Party Within Policy Environments. Bars show weighted mean mask usage (percentage of contacts with masks) for each party, faceted by county mask mandate policy. Error bars indicate standard errors."
#| fig-height: 8
#| fig-width: 6

# Masks by policy - faceted vertically
ggplot(behaviors_by_policy, aes(x = political_party, y = mask_mean, fill = political_party)) +
  geom_col(width = 0.7, alpha = 0.8, color = "black") +
  geom_errorbar(aes(ymin = mask_mean - mask_se, ymax = mask_mean + mask_se),
                width = 0.2) +
  geom_text(aes(label = sprintf("%.0f%%", mask_mean)), vjust = -1.2, size = 3.5, fontface = "bold") +
  facet_wrap(~ policy, ncol = 1) +
  scale_fill_manual(values = party_colors) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.2)), limits = c(0, 85)) +
  labs(x = NULL, y = "Mask Usage (%)") +
  theme_cowplot() +
  theme(
    legend.position = "none",
    axis.text.x = element_text(angle = 45, hjust = 1, size = 11),
    strip.text = element_text(face = "bold", size = 14),
    strip.background = element_blank()
  )
```

The results clearly demonstrate that partisan differences are not simply artifacts of differential policy exposure. Within each policy environment, Republicans reported more daily contacts than Democrats: the gap was 1.4 contacts in No Mandate areas, 1.0 contacts in Less Strict areas, and 1.3 contacts in Strict mandate areas. Similarly, Democrats reported higher mask usage than Republicans across all policy contexts: the gap was 9 percentage points in No Mandate areas, 11 percentage points in Less Strict areas, and 10 percentage points in Strict mandate areas.

These findings suggest that partisan identity shapes health behaviors above and beyond the effects of local policy. While stricter policies are associated with higher mask usage overall, the Republican-Democrat gap in mask-wearing remains remarkably stable across policy environments. This pattern is consistent with partisanship operating through mechanisms other than policy compliance---such as differential risk perception, social norms within partisan networks, or ideological commitments---that persist regardless of the local regulatory context.


<!-- === Appendix I: Geographic Variation in Incidence Responsiveness === -->

## APPENDIX I. Geographic Variation in Incidence Responsiveness {#sec-geographic-variation}

```{r}
#| label: load-wave-responsiveness
#| include: false

# Load wave-by-wave incidence responsiveness results
wave_resp <- read_csv("../../out/tables/wave_incidence_responsiveness.csv",
                      show_col_types = FALSE) %>%
  mutate(
    wave_label = case_when(
      wave == 1 ~ "Apr '20",
      wave == 2 ~ "Jun '20",
      wave == 3 ~ "Sep '20",
      wave == 4 ~ "Dec '20",
      wave == 5 ~ "Feb '21",
      wave == 6 ~ "May '21"
    ),
    wave_label = factor(wave_label, levels = c("Apr '20", "Jun '20", "Sep '20", "Dec '20", "Feb '21", "May '21"))
  )

# Summary statistics
n_waves_significant <- sum(wave_resp$gap_change < 0, na.rm = TRUE)
avg_gap_change <- mean(wave_resp$gap_change, na.rm = TRUE)
max_rep_reduction <- wave_resp %>% slice_min(effect_rep, n = 1)
min_rep_reduction <- wave_resp %>% slice_max(effect_rep, n = 1)
```

The main text presents models that pool across survey waves, using temporal variation in pandemic conditions (alongside geographic variation) to identify behavioral responses. To isolate the role of geographic variation, we estimate separate models for each survey wave. Within a single wave, variation in county-level incidence rates is purely geographic, reflecting cross-sectional differences in local pandemic severity rather than temporal trends. This approach allows us to examine whether partisan differences in responsiveness to local conditions are consistent across different periods of the pandemic.

@tbl-wave-responsiveness presents the estimated change in daily contacts associated with a 10% increase in local (log) incidence rates, by party and wave. All models include state fixed effects to control for time-invariant state characteristics. @fig-wave-responsiveness visualizes these results, showing how the incidence-responsiveness of each partisan group varied over the study period.

```{r}
#| label: tbl-wave-responsiveness
#| tbl-cap: "Contact Response to Local Incidence by Wave and Party"

wave_resp %>%
  select(wave_label, n, mean_inc_rate, effect_dem, effect_ind, effect_rep, gap_change) %>%
  kable(
    col.names = c("Wave", "N", "Mean Inc. Rate", "Democrat", "Independent", "Republican", "R-D Gap Change"),
    digits = c(0, 0, 1, 3, 3, 3, 3),
    align = c("l", "r", "r", "r", "r", "r", "r"),
    booktabs = TRUE
  ) %>%
  kable_styling(latex_options = c("hold_position", "scale_down")) %>%
  footnote(
    general = "Effects represent the change in daily contacts associated with a 10% increase in local log incidence rate. Negative values indicate contact reduction. R-D Gap Change shows how the Republican-Democrat contact gap changes (negative = gap narrows). All models include state fixed effects, city indicators, and interview day controls.",
    general_title = "Note: ",
    footnote_as_chunk = TRUE
  )
```

```{r}
#| label: fig-wave-responsiveness
#| fig-cap: "Contact Response to Local Incidence Across Survey Waves. Points show the estimated change in daily contacts associated with a 10% increase in local log incidence rates, by party and wave. Error bars indicate standard errors. Dashed line at zero represents no response."
#| fig-height: 5
#| fig-width: 9

# Party colors (matching other BICS plots)
party_colors <- c("Democrat" = "#326ba4", "Independent" = "#6db753", "Republican" = "#c50000")

# Reshape for plotting
wave_plot_data <- wave_resp %>%
  select(wave_label, effect_dem, effect_ind, effect_rep, se_dem, se_ind, se_rep) %>%
  pivot_longer(
    cols = starts_with("effect_"),
    names_to = "party",
    names_prefix = "effect_",
    values_to = "effect"
  ) %>%
  mutate(
    se = case_when(
      party == "dem" ~ wave_resp$se_dem[match(wave_label, wave_resp$wave_label)],
      party == "ind" ~ wave_resp$se_ind[match(wave_label, wave_resp$wave_label)],
      party == "rep" ~ wave_resp$se_rep[match(wave_label, wave_resp$wave_label)]
    ),
    party = case_when(
      party == "dem" ~ "Democrat",
      party == "ind" ~ "Independent",
      party == "rep" ~ "Republican"
    ),
    party = factor(party, levels = c("Republican", "Independent", "Democrat"))
  )

ggplot(wave_plot_data, aes(x = wave_label, y = effect, color = party, group = party)) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "gray50") +
  geom_line(linewidth = 0.8, alpha = 0.7) +
  geom_point(size = 3) +
  geom_errorbar(aes(ymin = effect - se, ymax = effect + se), width = 0.15) +
  scale_color_manual(values = party_colors) +
  labs(
    x = "Survey Wave",
    y = "Change in Daily Contacts\n(per 10% increase in log incidence)",
    color = "Party"
  ) +
  theme_cowplot() +
  theme(
    legend.position = "bottom",
    axis.text.x = element_text(angle = 45, hjust = 1)
  )
```

### Interpretation

The wave-by-wave analysis reveals several patterns. First, Republicans consistently show stronger contact reductions in response to local incidence than Democrats across most survey waves. In `r n_waves_significant` of 6 waves, the Republican-Democrat gap change is negative, indicating that higher local incidence narrows the partisan contact gap. Second, the magnitude of responsiveness varies over time, with the strongest Republican responses occurring during `r max_rep_reduction$wave_label` when mean incidence was `r sprintf("%.1f", max_rep_reduction$mean_inc_rate)` cases per 100,000.

These within-wave analyses provide complementary evidence to the pooled models in the main text. By isolating geographic variation in incidence, we demonstrate that partisan differences in responsiveness are not artifacts of differential exposure to temporal trends in the pandemic. Rather, when comparing individuals facing different local conditions at the same point in time, Republicans are more responsive to higher local incidence than Democrats---a pattern that helps explain the narrowing of partisan gaps during periods of elevated disease burden.


<!-- === Appendix K subsections: Incidence vs Mortality, Mask Incidence, Perception Response === -->

### Incidence vs. Mortality: Why We Focus on Mortality {#sec-incidence-vs-mortality}

The main text focuses on mortality rather than incidence when discussing partisan responses to pandemic conditions. This section explains that choice and presents the full comparison of incidence and mortality effects.

**Key finding: Mortality was the stronger predictor of contact behavior, while incidence was the stronger predictor of mask-wearing.**

@fig-pandemic-response-detail above displays the full comparison. For contacts (left column), mortality effects are larger than incidence effects for all partisan groups, and statistically significant for Republicans. For masks (right column), the pattern reverses: incidence effects are larger, particularly for Democrats.

This differential pattern has a plausible interpretation:

1. **Mortality and contacts**: Deaths are concrete, visible, and emotionally salient. They are reported in local news, experienced through social networks, and marked by funerals and obituaries. When deaths rise, the threat becomes tangible—prompting reductions in social contact. Republicans showed the largest mortality response, consistent with the main text's finding that their contact behavior swung more dramatically with conditions.

2. **Incidence and masks**: Case counts, by contrast, are more abstract and mediated by testing availability and reporting. However, rising case counts may trigger mask mandates, workplace policies, and social pressure to mask—particularly in areas where cases are visibly surging. Democrats, who were more likely to live in areas with mask mandates and to follow public health guidance, showed the largest incidence response for masks.

@fig-gap-changes summarizes the implications for polarization: mortality narrows the contact gap (Republicans reduce contacts more), while incidence slightly widens the mask gap (Democrats increase masking more). This supports the main text's focus on mortality for the contact behavior analysis, where the key finding is that objective threat can reduce polarization—at least temporarily.

The sections below provide additional detail on mask responses to incidence (@sec-mask-incidence-response) and perception responses to both conditions (@sec-perception-response).

### Mask Usage Response to Incidence {#sec-mask-incidence-response}

@fig-incidence-mask-response parallels @fig-mortality-response in the main text but shows mask usage response to incidence rather than contacts response to mortality. This figure illustrates the finding that incidence was the stronger predictor of mask-wearing, particularly for Democrats.

```{r}
#| label: fig-incidence-mask-response
#| fig-cap: "Mask Usage Response to Incidence. Panel A shows predicted mask usage (percentage of contacts with masks) by party at baseline (average incidence) and when incidence is one standard deviation higher. Panel B shows the Democrat-Republican gap in mask usage at baseline and at higher incidence. All partisan groups increase mask usage when incidence rises, with Democrats showing the largest response."
#| fig-height: 5
#| fig-width: 10

library(patchwork)

# Calculate predicted mask usage at baseline and +1 SD incidence
# Baseline = raw weighted means; +1 SD = baseline + incidence coefficient
pred_masks <- tibble(
  party = rep(c("Republican", "Independent", "Democrat"), 2),
  condition = rep(c("Baseline", "+1 SD Incidence"), each = 3),
  mask_usage = c(
    # Baseline (raw means)
    mask_rep, mask_ind, mask_dem,
    # +1 SD incidence (baseline + effect)
    mask_rep + masks_inc_effect_rep,
    mask_ind + masks_inc_effect_ind,
    mask_dem + masks_inc_effect_dem
  )
) %>%
  mutate(
    party = factor(party, levels = c("Republican", "Independent", "Democrat")),
    condition = factor(condition, levels = c("Baseline", "+1 SD Incidence"))
  )

# Calculate gaps (D-R since Democrats mask more)
gap_baseline <- mask_dem - mask_rep
gap_high_inc <- (mask_dem + masks_inc_effect_dem) - (mask_rep + masks_inc_effect_rep)

gap_data <- tibble(
  condition = factor(c("Baseline", "+1 SD Incidence"), levels = c("Baseline", "+1 SD Incidence")),
  gap = c(gap_baseline, gap_high_inc)
)

# Panel A: Predicted mask usage by party
p_masks <- ggplot(pred_masks, aes(x = condition, y = mask_usage, fill = party)) +
  geom_col(position = position_dodge(width = 0.8), width = 0.7, alpha = 0.8, color = "black") +
  geom_text(aes(label = sprintf("%.1f", mask_usage), group = party),
            position = position_dodge(width = 0.8), vjust = -0.5, size = 3.5, fontface = "bold") +
  scale_fill_manual(values = party_colors) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.15))) +
  labs(x = NULL, y = "Mask Usage (%)", title = "A. Predicted Mask Usage by Party", fill = "Party") +
  theme_cowplot() +
  theme(
    plot.title = element_text(face = "bold", size = 14),
    legend.position = "bottom"
  )

# Panel B: D-R Gap
p_gap <- ggplot(gap_data, aes(x = condition, y = gap)) +
  geom_col(width = 0.6, fill = "#1b9e77", alpha = 0.8, color = "black") +
  geom_text(aes(label = sprintf("%.1f", gap)), vjust = -0.5, size = 4, fontface = "bold") +
  geom_hline(yintercept = 0, linetype = "dashed", color = "gray50") +
  scale_y_continuous(expand = expansion(mult = c(0.05, 0.15))) +
  labs(x = NULL, y = "D-R Gap (pp)", title = "B. Democrat-Republican Gap") +
  theme_cowplot() +
  theme(
    plot.title = element_text(face = "bold", size = 14)
  )

# Combine
p_masks + p_gap + plot_layout(widths = c(1.5, 1))
```

### Perception Response to Pandemic Conditions {#sec-perception-response}

@fig-perception-response shows how COVID-19 concern (the percentage reporting being "very concerned") responded to prior week's pandemic conditions by party. This parallels @fig-pandemic-response-detail but examines perceptions rather than behaviors.

```{r}
#| label: fig-perception-response
#| fig-cap: "Risk Perception Response to Prior Week's Pandemic Conditions by Party. Left column shows response to incidence; right column shows response to mortality. Bars show the change in the percentage reporting being 'very concerned' about COVID-19 associated with a one standard deviation increase in log incidence or log mortality. Models include county fixed effects. Error bars indicate standard errors."
#| fig-height: 5
#| fig-width: 10

library(patchwork)

# Create data frame for perception response plotting
perception_response_data <- tibble(
  party = rep(c("Democrat", "Independent", "Republican"), 2),
  condition = rep(c("Incidence", "Mortality"), each = 3),
  effect = c(
    concern_inc_effect_dem, concern_inc_effect_ind, concern_inc_effect_rep,
    concern_mort_effect_dem, concern_mort_effect_ind, concern_mort_effect_rep
  ),
  se = c(
    concern_inc_se_dem, concern_inc_se_ind, concern_inc_se_rep,
    concern_mort_se_dem, concern_mort_se_ind, concern_mort_se_rep
  )
) %>%
  mutate(
    party = factor(party, levels = c("Republican", "Independent", "Democrat")),
    condition = factor(condition, levels = c("Incidence", "Mortality"))
  )

# Calculate y-axis limits
y_range <- perception_response_data %>%
  summarise(ymin = min(effect - se, na.rm = TRUE), ymax = max(effect + se, na.rm = TRUE))
y_lim <- c(min(y_range$ymin, -1), max(y_range$ymax, 1)) * 1.15

# Create the plot
ggplot(perception_response_data, aes(x = party, y = effect, fill = party)) +
  geom_col(alpha = 0.7, width = 0.7, color = "black") +
  geom_errorbar(aes(ymin = effect - se, ymax = effect + se), width = 0.2, color = "gray30") +
  geom_hline(yintercept = 0, linetype = "dashed", color = "gray50") +
  facet_wrap(~ condition, ncol = 2) +
  scale_fill_manual(values = party_colors) +
  scale_y_continuous(limits = y_lim) +
  labs(x = NULL, y = "Change in 'Very Concerned' (pp) per 1 SD") +
  theme_cowplot() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    strip.text = element_text(face = "bold", size = 13),
    strip.background = element_blank(),
    legend.position = "none",
    panel.border = element_rect(color = "gray70", fill = NA, linewidth = 0.5)
  )
```

Unlike health behaviors, which showed clear differential responses to pandemic conditions (mortality predicting contacts, incidence predicting masks), risk perception showed weaker and more inconsistent relationships with local pandemic conditions. This suggests that partisan differences in concern were more stable and less responsive to objective signals about pandemic severity---consistent with the finding in the main text that perception gaps remained relatively constant even as behavior gaps fluctuated.


<!-- === Appendix L subsection: Metro Detail === -->

### Urban-Rural Behavior Levels by Party {#sec-metro-detail}

@fig-metro-gaps displays the Republican-Democrat gap in contacts and mask-wearing by metropolitan status. The gaps are smaller in non-metropolitan areas for both behaviors: the contact gap drops from 1.3 to 0.3 additional daily contacts, and the mask gap drops from 9 to 7 percentage points. This pattern helps explain why the South---which has a higher proportion of non-metropolitan residents---showed smaller partisan gaps in the main analysis.

```{r}
#| label: fig-metro-gaps
#| fig-cap: "Republican-Democrat Gaps by Metropolitan Status. Each panel compares the partisan gap in metropolitan versus non-metropolitan areas. Panel A shows the contact gap (R−D) and Panel B shows the mask gap (D−R). Estimates are adjusted for survey wave and local mortality rates."
#| fig-height: 4
#| fig-width: 7

library(cowplot)
library(patchwork)

# Load adjusted metro data
behaviors_by_metro <- read_csv("../../out/tables/adj_behaviors_by_metro.csv", show_col_types = FALSE)

# Calculate R-D gaps by metro status
gaps_by_metro <- behaviors_by_metro %>%
  select(metro_status, political_party, contacts_adj_mean, mask_adj_mean) %>%
  pivot_wider(names_from = political_party,
              values_from = c(contacts_adj_mean, mask_adj_mean)) %>%
  mutate(
    contacts_gap_rd = contacts_adj_mean_Republican - contacts_adj_mean_Democrat,
    mask_gap_dr = mask_adj_mean_Democrat - mask_adj_mean_Republican,
    metro_status = factor(metro_status, levels = c("Metropolitan", "Non-Metropolitan"))
  )

# Colors for metro status
metro_colors <- c("Metropolitan" = "#4a4a4a", "Non-Metropolitan" = "#a0a0a0")

# Contact gap plot (R-D, so positive = R > D)
p_contact_gap <- ggplot(gaps_by_metro, aes(x = metro_status, y = contacts_gap_rd, fill = metro_status)) +
  geom_col(width = 0.6, color = "black") +
  geom_text(aes(label = sprintf("%.1f", contacts_gap_rd)), vjust = -0.5, size = 4, fontface = "bold") +
  geom_hline(yintercept = 0, linetype = "dashed", color = "gray50") +
  scale_fill_manual(values = metro_colors) +
  scale_y_continuous(expand = expansion(mult = c(0.05, 0.15))) +
  labs(x = NULL, y = "R−D Gap (contacts)", title = "A. Contact Gap") +
  theme_cowplot() +
  theme(
    plot.title = element_text(face = "bold", size = 11),
    legend.position = "none"
  )

# Mask gap plot (D-R, so positive = D > R)
p_mask_gap <- ggplot(gaps_by_metro, aes(x = metro_status, y = mask_gap_dr, fill = metro_status)) +
  geom_col(width = 0.6, color = "black") +
  geom_text(aes(label = sprintf("%.0f", mask_gap_dr)), vjust = -0.5, size = 4, fontface = "bold") +
  geom_hline(yintercept = 0, linetype = "dashed", color = "gray50") +
  scale_fill_manual(values = metro_colors) +
  scale_y_continuous(expand = expansion(mult = c(0.05, 0.15))) +
  labs(x = NULL, y = "D−R Gap (pp)", title = "B. Mask Gap") +
  theme_cowplot() +
  theme(
    plot.title = element_text(face = "bold", size = 11),
    legend.position = "none"
  )

# Combine horizontally
p_contact_gap + p_mask_gap
```

@fig-metro-behaviors-detail displays health behavior levels by party and metropolitan status, showing that all partisan groups were less cautious in non-metropolitan areas.

```{r}
#| label: fig-metro-behaviors-detail
#| fig-cap: "Adjusted Health Behaviors by Party and Metropolitan Status. Each panel shows behavior levels for the three partisan groups, comparing metropolitan (darker bars) versus non-metropolitan (lighter bars) areas. Panel A shows daily contacts, Panel B shows mask usage, and Panel C shows vaccine acceptance rates (Wave 6 only). Estimates are adjusted for survey wave and local mortality rates. Error bars indicate standard errors."
#| fig-height: 9
#| fig-width: 6

library(cowplot)
library(patchwork)

# Load adjusted metro data (emmeans with wave + mortality controls)
behaviors_by_metro <- read_csv("../../out/tables/adj_behaviors_by_metro.csv", show_col_types = FALSE) %>%
  mutate(
    political_party = factor(political_party, levels = c("Republican", "Independent", "Democrat")),
    metro_status = factor(metro_status, levels = c("Metropolitan", "Non-Metropolitan")),
    contacts_mean = contacts_adj_mean,
    contacts_se = contacts_adj_se,
    mask_mean = mask_adj_mean,
    mask_se = mask_adj_se,
    vax_mean = vax_adj_mean,
    vax_se = vax_adj_se
  )

# Party colors with metro/non-metro shading
party_colors_metro <- c(
  "Republican_Metropolitan" = "#c50000",
  "Republican_Non-Metropolitan" = "#ff6666",
  "Independent_Metropolitan" = "#6db753",
  "Independent_Non-Metropolitan" = "#a8d99a",
  "Democrat_Metropolitan" = "#326ba4",
  "Democrat_Non-Metropolitan" = "#7eb3e0"
)

behaviors_by_metro <- behaviors_by_metro %>%
  mutate(party_metro = paste(political_party, metro_status, sep = "_"))

# Contacts plot
p_contacts <- ggplot(behaviors_by_metro, aes(x = political_party, y = contacts_mean, fill = party_metro)) +
  geom_col(position = position_dodge(width = 0.8), width = 0.7, color = "black") +
  geom_errorbar(aes(ymin = contacts_mean - contacts_se, ymax = contacts_mean + contacts_se),
                position = position_dodge(width = 0.8), width = 0.2) +
  scale_fill_manual(values = party_colors_metro) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.1))) +
  labs(x = NULL, y = "Daily Contacts", title = "A. Contacts", fill = NULL) +
  theme_cowplot() +
  theme(plot.title = element_text(face = "bold", size = 11), legend.position = "none")

# Masks plot
p_masks <- ggplot(behaviors_by_metro, aes(x = political_party, y = mask_mean, fill = party_metro)) +
  geom_col(position = position_dodge(width = 0.8), width = 0.7, color = "black") +
  geom_errorbar(aes(ymin = mask_mean - mask_se, ymax = mask_mean + mask_se),
                position = position_dodge(width = 0.8), width = 0.2) +
  scale_fill_manual(values = party_colors_metro) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.1)), limits = c(0, 80)) +
  labs(x = NULL, y = "Mask Usage (%)", title = "B. Mask Usage", fill = NULL) +
  theme_cowplot() +
  theme(plot.title = element_text(face = "bold", size = 11), legend.position = "none")

# Vaccine acceptance plot
p_vax <- ggplot(behaviors_by_metro, aes(x = political_party, y = vax_mean, fill = party_metro)) +
  geom_col(position = position_dodge(width = 0.8), width = 0.7, color = "black") +
  geom_errorbar(aes(ymin = vax_mean - vax_se, ymax = vax_mean + vax_se),
                position = position_dodge(width = 0.8), width = 0.2) +
  scale_fill_manual(values = party_colors_metro,
                    labels = c("Democrat_Metropolitan" = "Metro", "Democrat_Non-Metropolitan" = "Non-Metro",
                              "Independent_Metropolitan" = "Metro", "Independent_Non-Metropolitan" = "Non-Metro",
                              "Republican_Metropolitan" = "Metro", "Republican_Non-Metropolitan" = "Non-Metro")) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.1)), limits = c(0, 100)) +
  labs(x = NULL, y = "Vaccine Acceptance (%)", title = "C. Vaccine Acceptance", fill = "Metro Status") +
  theme_cowplot() +
  theme(plot.title = element_text(face = "bold", size = 11), legend.position = "bottom") +
  guides(fill = guide_legend(nrow = 1, override.aes = list(fill = c("#326ba4", "#7eb3e0", "#6db753", "#a8d99a", "#c50000", "#ff6666"))))

# Combine vertically
p_contacts / p_masks / p_vax
```

The figure reveals that the urban-rural shift toward less protective behavior is consistent across all three partisan groups. Notably, Independents show the largest urban-rural differences, particularly for mask-wearing and vaccine acceptance.


<!-- === Appendix L subsections: Division Maps + State Maps === -->

### Partisan Gaps by Census Division {#sec-division-maps}

@fig-division-gap-map displays partisan gaps at the census division level (9 divisions), providing more geographic granularity than the 4 census regions. **Note**: The East South Central division (AL, KY, MS, TN) shows a reversed pattern with Democrats reporting more contacts and higher mask usage than Republicans. This anomaly likely reflects sampling variability given the smaller sample sizes in that division (n=166 Democrats, n=205 Republicans).

```{r}
#| label: fig-division-gap-map
#| fig-cap: "Geographic Distribution of Adjusted Partisan Gaps by Census Division. Panel A shows the Republican-Democrat gap in daily contacts (purple = Republicans report more contacts; teal = Democrats report more). Panel B shows the Democrat-Republican gap in mask usage (gold = Democrats mask more; teal = Republicans mask more). Estimates are adjusted for survey wave and local mortality rates. White indicates no gap. Black borders indicate census division boundaries."
#| fig-height: 8
#| fig-width: 8

library(usmap)
library(patchwork)
library(sf)

# Load division gaps - using adjusted estimates
gaps_by_division <- read_csv("../../out/tables/adj_gaps_by_division.csv", show_col_types = FALSE)

# Create division-to-state mapping
division_map <- tibble(
  state = c(
    # New England
    "Connecticut", "Maine", "Massachusetts", "New Hampshire", "Rhode Island", "Vermont",
    # Middle Atlantic
    "New Jersey", "New York", "Pennsylvania",
    # East North Central
    "Illinois", "Indiana", "Michigan", "Ohio", "Wisconsin",
    # West North Central
    "Iowa", "Kansas", "Minnesota", "Missouri", "Nebraska", "North Dakota", "South Dakota",
    # South Atlantic
    "Delaware", "District of Columbia", "Florida", "Georgia", "Maryland",
    "North Carolina", "South Carolina", "Virginia", "West Virginia",
    # East South Central
    "Alabama", "Kentucky", "Mississippi", "Tennessee",
    # West South Central
    "Arkansas", "Louisiana", "Oklahoma", "Texas",
    # Mountain
    "Arizona", "Colorado", "Idaho", "Montana", "Nevada", "New Mexico", "Utah", "Wyoming",
    # Pacific
    "Alaska", "California", "Hawaii", "Oregon", "Washington"
  ),
  census_division = c(
    rep("New England", 6), rep("Middle Atlantic", 3),
    rep("East North Central", 5), rep("West North Central", 7),
    rep("South Atlantic", 9), rep("East South Central", 4), rep("West South Central", 4),
    rep("Mountain", 8), rep("Pacific", 5)
  )
)

# Merge with gaps data
division_gaps_for_map <- division_map %>%
  left_join(gaps_by_division, by = "census_division")

# Get state polygons and create division boundaries
us_states_sf <- us_map(regions = "states") %>%
  left_join(division_map, by = c("full" = "state"))

division_boundaries <- us_states_sf %>%
  group_by(census_division) %>%
  summarise(geometry = st_union(geom), .groups = "drop")

# Contact gap map - diverging scale: purple (R>D) vs teal (D>R)
p_contact_map <- plot_usmap(data = division_gaps_for_map, values = "contacts_gap_rd", color = "white", linewidth = 0.2) +
  geom_sf(data = division_boundaries, fill = NA, color = "black", linewidth = 0.6, inherit.aes = FALSE) +
  scale_fill_gradient2(
    low = "#1b9e77", mid = "white", high = "#5c2d91",
    midpoint = 0,
    name = "R−D Gap\n(contacts)",
    na.value = "gray90"
  ) +
  labs(title = "A. Contact Gap (Republican − Democrat)") +
  theme(
    plot.title = element_text(face = "bold", size = 12, hjust = 0.5),
    legend.position = "right"
  )

# Mask gap map - diverging scale: gold (D>R) vs teal (R>D)
p_mask_map <- plot_usmap(data = division_gaps_for_map, values = "mask_gap_dr", color = "white", linewidth = 0.2) +
  geom_sf(data = division_boundaries, fill = NA, color = "black", linewidth = 0.6, inherit.aes = FALSE) +
  scale_fill_gradient2(
    low = "#1b9e77", mid = "white", high = "#b8860b",
    midpoint = 0,
    name = "D−R Gap\n(pp)",
    na.value = "gray90"
  ) +
  labs(title = "B. Mask Gap (Democrat − Republican)") +
  theme(
    plot.title = element_text(face = "bold", size = 12, hjust = 0.5),
    legend.position = "right"
  )

# Stack vertically
p_contact_map / p_mask_map
```

### State-Level Partisan Gaps {#sec-state-maps}

@fig-state-maps provides a more granular view, mapping state-level partisan differences in contacts and mask usage. **Important caveat**: The BICS sample was designed to be nationally representative, not representative at the state level. State-level estimates should therefore be interpreted with caution, as they may reflect sampling variability rather than true geographic differences. We present these maps for exploratory purposes only; the regional and division-level analyses above provide more reliable geographic comparisons. States with fewer than 10 respondents per party are excluded.

```{r}
#| label: fig-state-maps
#| fig-cap: "State-Level Adjusted Partisan Gaps in Health Behaviors. Panel A shows the Republican-Democrat gap in daily contacts (purple = Republicans report more; teal = Democrats report more). Panel B shows the Democrat-Republican gap in mask usage (gold = Democrats mask more; teal = Republicans mask more). Estimates are adjusted for survey wave and local mortality rates. White indicates no gap; gray states have insufficient sample size."
#| fig-height: 8
#| fig-width: 10

library(usmap)
library(patchwork)

# Load state-level gaps - using adjusted estimates
gaps_by_state <- read_csv("../../out/tables/adj_gaps_by_state.csv", show_col_types = FALSE) %>%
  mutate(state = STATE)  # usmap needs 'state' column with full names

# Contact gap map - diverging: purple (R>D) vs teal (D>R)
p_contacts_map <- plot_usmap(data = gaps_by_state, values = "contacts_gap_rd", color = "gray50") +
  scale_fill_gradient2(
    low = "#1b9e77", mid = "white", high = "#5c2d91",
    midpoint = 0,
    name = "R−D Gap\n(contacts)",
    na.value = "gray90",
    limits = c(-5, 7)
  ) +
  labs(title = "A. Contact Gap (Republican − Democrat)") +
  theme(
    plot.title = element_text(face = "bold", size = 12, hjust = 0.5),
    legend.position = "right"
  )

# Mask gap map - diverging: gold (D>R) vs teal (R>D)
p_masks_map <- plot_usmap(data = gaps_by_state, values = "mask_gap_dr", color = "gray50") +
  scale_fill_gradient2(
    low = "#1b9e77", mid = "white", high = "#b8860b",
    midpoint = 0,
    name = "D−R Gap\n(pp)",
    na.value = "gray90",
    limits = c(-10, 35)
  ) +
  labs(title = "B. Mask Gap (Democrat − Republican)") +
  theme(
    plot.title = element_text(face = "bold", size = 12, hjust = 0.5),
    legend.position = "right"
  )

# Combine vertically
p_contacts_map / p_masks_map
```

The maps reveal substantial heterogeneity in partisan gaps across states. For contact behavior, most states show the expected pattern of Republicans reporting more contacts than Democrats (red shading), but the magnitude varies considerably---from nearly no gap in some Southern states to gaps exceeding 5 contacts per day in states like Michigan and Colorado. A few states show reversed patterns (blue shading) where Democrats report more contacts, though these often reflect small sample sizes.

For mask usage, the Democrat-Republican gap (blue shading indicates Democrats masking more) is positive in nearly all states, but again varies substantially in magnitude. Western states tend to show larger mask gaps, while some Southern states show smaller differences---consistent with the regional patterns documented in the main text.


